---
title: "Lesson 4 Challenge problems"
author: "Gabrielle Ferra"
date: '2022-08-07'
output: html_document
---

If you already downloaded the dataset we worked with in the "Lesson4_practice.Rmd" notebook, run this code chunk:

```{r}
clinical <- read.csv("data/clinical.csv")      
```

If you have not already downloaded the dataset from the "Lesson4_practice.Rmd" notebook, run this code chunk:

```{r}
dir.create("data") 
 #Download the file here: https://github.com/gferra97/Coding_for_Cancer_2023/blob/3bbd87c817d750e36d83bd8558347972baa91bf9/clinical.csv

clinical <- read.csv("data/clinical.csv")      
```

Question 1: 

What is the difference in results between clinical[1] and clinical[ , 1] (using our clinical data frame)?

```{r}
# your code here..

```

Question 2:
a)	Code as many different ways possible to extract the column “pack_years_smoked”
b)	Extract the first 6 rows for only “sample_type”, “ajcc_pathologic_stage”, and “age_at_diagnosis”.
c)	Calculate the range and mean for cigarettes_per_day.
d)	What age is the oldest patient (at diagnosis) in years? Youngest? (Note: age_at_diagnosis is given in days)

```{r}
# your code here..

```
