---
title: "Lesson 5a Challenge problems"
author: "Magdalena Russell & Katherine Brower"
date: '2023-08-08'
output: html_document
---

All exercises use the clinical data object we made earlier:

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)

# Load the data
clinical <- read.csv("data/clinical.csv")
```

Create a new object from clinical called demographics_BLCA that includes only the height, bmi, gender, ethnicity, and race.

```{r}
# your code here...

```

Code as many ways to select the columns of barcode, project_id, tissue_or_organ_of_origin, primary_diagnosis, and sample_type from the clinical object. Hint there are 6 ways!

```{r}
# your code here...

```

Create a new object from clinical called demographics_BLCA_current_smokers that includes only the height, bmi, gender, ethnicity, and race. However, restrict data such that only current smokers are included (hint: relevant column is called Tobacco.smoking.history).

```{r}
# your code here...

```


Create a new object from clinical called demographics_BCLA_mutation_count that includes only the height, bmi, gender, ethnicity, race, and count.ERCC2.signature.mutations. However, restrict the data that only patients with presence of prostate cancer are included (hint: relevant column is called Presence.of.prostate.cancer)

Once you have made the dataset, calculate the mean count of ERCC2 signature mutations of the patients (hint: relavent column is count.ERCC2.signature.mutations)
``` {r}
#your code here...

#calculate the mean count of ERCC2 signature mutations:
```
