---
title: "Lesson 5a Challenge problems"
author: "Magdalena Russell & Katherine Brower"
date: '2023-08-08'
output: html_document
---

All exercises use the clinical object we created in the readings:

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)

# Load the data
clinical <- read.csv("data/clinical.csv")
```

Create a new object from clinical called demographics_BLCA that includes only the height, bmi, gender, ethnicity, and race.

```{r}
#creating the new object using the select() function
demographics_BLCA <- select(clinical, height, bmi, gender, ethnicity, race)
```

Code as many ways to select the columns of barcode, project_id, tissue_or_organ_of_origin, primary_diagnosis, and sample_type from the clinical object. Hint there are 6 ways!

```{r}
#6 ways to create the same object with the same selected columns!
try_1 <- select(clinical, barcode, project_id, tissue_or_organ_of_origin, primary_diagnosis, sample_type)
try_2 <- select(clinical, c(1, 2, 3, 4, 5))
try_3 <- select(clinical, 1:5)
try_4 <- clinical[ , 1:5]
try_5 <- clinical[ , c(1, 2, 3, 4, 5)]
try_6 <- clinical[c("barcode", "project_id", "tissue_or_organ_of_origin", "primary_diagnosis", "sample_type")]
```

Create a new object from clinical called demographics_BLCA_current_smokers that includes only the height, bmi, gender, ethnicity, and race. However, restrict data such that only current smokers are included (hint: relevant column is called Tobacco.smoking.history).

```{r}
# creating the new object with the conditions given in the question
demographics_BLCA_current_smokers <- select(filter(clinical, Tobacco.smoking.history == "Current smoker"), height, bmi, gender, ethnicity, race)

#calculate the mean of the count of the ERCC2 signature mutations. Remember that we can use $ operator with data frames! (Hint: the column we are looking for is count.ERCC2.signature.mutations)
mean(demographics_BCLA_mutation_count$count.ERCC2.signature.mutations, na.rm = TRUE)
#the answer is 64.31059

```


