---
title: "Lesson6_key"
author: "Sam Hart & Katheirne Brower"
date: "2023-08-10"
output: html_document
---

__Under each heading below, type your code and run it. Feel free to add notes and comments to help you learn.__ 


# load library/package
```{r}
library(tidyverse)
```

# import data

```{r}
clinical <- read_csv("data/clinical.csv")
```

# show categories in gender

```{r}
unique(clinical$gender)
```

# count number of individuals of each gender

```{r}
gender_count <- clinical %>%
  group_by(gender) %>%
  tally() 

View(gender_count)
```

# summarize average birth year by gender

```{r}
avg_YOB_by_gender <- clinical %>%
  group_by(gender) %>%
  summarize(mean_year_of_birth = mean(year_of_birth, na.rm = TRUE))

# optional:
avg_YOB_by_gender_2 <- clinical %>%
  group_by(gender) %>%
  summarize(mean_age = 2023 - mean(year_of_birth, na.rm = TRUE))

View(avg_YOB_by_gender)
```

# summarize average age at diagnosis (in days) by gender

```{r}
avg_age_days_by_gender <- clinical %>%
  group_by(gender) %>%
  summarize(mean_age_at_diagnosis_days = mean(age_at_diagnosis, na.rm = TRUE))
```

# summarize average, median, minimum, and maximum 
# age at diagnosis (in days) by gender

```{r}
clinical_summarized_by_gender <- clinical %>%
  group_by(gender) %>%
  summarize(a = mean(age_at_diagnosis, na.rm = TRUE),
            b = median(age_at_diagnosis, na.rm = TRUE),
            c = max(age_at_diagnosis, na.rm = TRUE),
            d = min(age_at_diagnosis, na.rm = TRUE))

View(clinical_summarized_by_gender)
```

# save clinical_summarized as a .csv file in our data directory

```{r}
write_csv(clinical_summarized_by_gender, "data/clinical_summarized.csv")
```

# counting number of distinct diagnoses
```{r}
diagnosis_counts <- clinical %>%
  count(primary_diagnosis) %>%
  arrange(n) 

```

# get names of cancer diagnoses with more than 1 patient
```{r}
specific_diagnoses <- diagnosis_counts %>%
  filter(n > 1)

```

# extract data specifying primary diagnoses to keep

```{r}
# One way
clinical_reduced2 <- clinical %>%
  filter(primary_diagnosis == "Papillary transitional cell carcinoma" | # OR
         primary_diagnosis == "Transitional cell carcinoma")

# Quicker way - less typing
clinical_reduced <- clinical %>%
  filter(primary_diagnosis %in% specific_diagnoses$primary_diagnosis)



```

# confirm diagnoses present are the ones we want

```{r}
clinical_reduced$primary_diagnosis %>% 
  unique()
```

# same as above but without an intermediate object

```{r}
clinical_reduced3 <- clinical %>%
  filter(primary_diagnosis %in% c("Transitional cell carcinoma", "Papillary transitional cell carcinoma"))
```
