---
title: "1.0_intro_to_R"
output: html_document
date: "2024-01-12"
---

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: 
Today's date:
More notes:


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}
#insert code here
```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk.

We are going to use these Rmd files to save our code from our activities in swirl(). 
You can write yourself notes in these files, these are meant for your and your learning!
You won't be able to run any code in the code chunks with swirl() open, so do any troubleshooting 
in swirl() before writing it down in here. 

Once you close out of swirl() you can run your code. 

Work through each activity in swirl(), but make sure to copy over code so you can keep 
a record of all the amazing code you write!


#add 3 and 4 together, using the '+' operator, just like in math class
```{r}

```

#try multiplying 5 by 6, using the * operator 
```{r}

```

#assign the value 10 to the object weight_kg, by typing weight_kg <- 10
```{r}

```

#print the variable you created above
```{r}

```

#convert kg to lb by multiplying weight_kg by 2.2
```{r}

```

#assign a new object called weight_lb with the value you got above
```{r}

```

#print your new object
```{r}

```

#re-assign weight_kg to 100
```{r}

```

#print weight_kg to check
```{r}

```

Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  
Once you exit swirl(), try running your code. 

