---
title: "Lesson 6 Challenge problems"
author: "Sam Hart & Katherine Brower"
date: "2023-08-10"
output: html_document
---

All exercises use the clinical object we created in the readings:

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)

# Load the data
clinical <- read.csv("data/clinical.csv")
```

Question 1:
A tibble called sample contains the following columns with categorical data (distinct categories are also specified):

    column_tissues with categories “breast”, “prostate”, “lung”, “bladder”, “colon”
    column_cancer with categories “adenocarcinoma”, “squamous cell carcinoma”, “DCIS”
    column_p53_mutation with categories “yes”, “no”

Which of the following commands are equivalent to 

    sample %>% 
      filter((column_tissues == "breast") | (column_tissues == "lung")) %>% 
      filter((column_cancer == "DCIS") | (column_cancer == "adenocarcinoma")) %>% 
      filter(column_p53_mutation != "yes")

Select all that apply.

    a)	sample %>% 
      filter(column_tissues %in% c("breast", "lung")) %>% 
      filter(column_cancer %in% c("DCIS", "adenocarcinoma")) %>% 
      filter(column_p53_mutation == "yes")
    b)	sample %>% 
      filter(column_tissues %in% c("breast", "lung")) %>% 
      filter(column_cancer %in% c("DCIS", "adenocarcinoma")) %>% 
      filter(column_p53_mutation %in% c("no"))
    c)	sample %>% 
      filter(column_tissues %in% c("breast", "lung"),
               column_cancer %in% c("DCIS", "adenocarcinoma"),
               column_p53_mutation == "no")
           
# The above question is conceptual - no code needed!
           
           
Question 2:
Create an object called smoke_complete from clinical that contains no missing data for cigarettes per day or age at diagnosis.
```{r}

```

How do you save the resulting table to file? How would you find this answer? Save the file as Lesson_6_Challenge_2.csv
```{r}

```

Question 3:
Create a new object called birth_complete that contains no missing data for year of birth or vital status.
```{r}

```

Question 4:
Extract all tumor stages with more than 100 cases (Hint: also check to see if there are any other missing/ambiguous data!)
```{r}

```

Question 5:
How many hispanic or latino individuals in clinical are not also white?
```{r}

```

