# 1.0 Intro to R

The gray box below is called a **code chunk**. We use it to write code.

```{r}

```

To run your code chunk, you can the green play button (Run current chunk) on the top right corner of the chunk. Let's try it here:

```{r}
# We use the "#" symbol to specify a line that R will not run as code. You can use it to write notes! Example below:

# This line of code prints the phrase Hello word
print("Hello World")
```


The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```


## Using R as a calculator

```{r}
# Add two values
12 + 4 
```

```{r}
5 + 4 * 6 - 2
```

```{r}
(5 + 4) * 6 - 2
```

```{r}
#Add your own equation for R to compute


```

## Assigning variables

We use the `<-` operator to store a value to a variable.

```{r}
x <- 15
```

To see what `x` is, we call it by typing it out:

```{r}
x
```

Alternatively, we can look in the `Environment` pane to see all objects and its values.

Here is an example for assigning objects and using the objects to form a new object:

```{r}
apples <- 5
oranges <- 6

total_fruits <- apples + oranges
total_fruits
```
