# 4.0 Functions in R

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```


## Functions

**Functions** are commands that allow us to do repeat complex tasks or actions. 

Let take this vector as our starting example:
```{r}
ages <- c(10,25,20,14,60,20)
ages
```

We can use the function `length` to find out the length of the vector:
```{r}
length(ages)
```

We can use the function `mean` to calculate the mean:
```{r}
mean(ages)
```

## More example

Let's look at other different functions:

- `abs()` take the absolute value of the input
```{r}
abs(-19.52)
```


- `floor()` rounds values **DOWN** to the nearest integer and `ceiling()` and rounds values **UP**.
```{r}
floor(5.340194)
```

```{r}
ceiling(5.340194)
```

- If you want to round to a specific decimal point, you can use `round()` and specify the number of `digits` like this:
```{r}
round(5.340194, digits=2)
```