---
title: "6.0 Data Frames worksheet"
author: "Katherine Brower (kbrower)"
date: "2024-06-11"
output: html_document
---
# 6.0 Data Frames

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Data Frames
We can use data frames to collect different variables we want to work with

Common dataframes are Excel spreadsheets and can contain different data types

We can create a dataframe by: 

df <- data.frame(title = variable1, title2 = variable2, title3 = variable3)

We can also view our dataframe using the function:
View(df)
```{r}
#create a data frame containing 
# - favorite_movies
# - favorite_animals
# - movie_length

favorite_movies <- c("Inside Out", "Howl's Moving Castle", "Nimona", "Guardian's of the Galaxy")

favorite_animals <- c("whaleshark", "puffin", "seahorse", "tiger")

movie_length <- c(95, 129, 98, 122)

# combine these variables into a single dataframe

favorites_df <- data.frame(movies = favorite_movies, animals = favorite_animals, movie_time = movie_length)

# let's look at our created dataframe

favorites_df
```

Each column in a data frame contains a variable:
-- Example: animals contains favorite_animals

If we want to look at a column individually, we can use the **$** operator

If we want to look at an individual value, we can use [] (the same way as any other vector)

```{r}
# access the whole column animals
favorites_df$animals
```

```{r}
# access only the value for whaleshark, first value

favorites_df$animals[1]
```

## Importing Data - Penguins
We'll be using the palmerpenguins package dataframe to do more practice!

We can download packages using the install.packages("package_name"). 
Once you download the package once, you don't need to download it again.

and load packages using: library(package_name)

```{r}
#install the palmerpenguins package
install.packages("palmerpenguins")
```

```{r}
#load the palmerpenguins package
library(palmerpenguins)
```
Let's look at the dataframe we just downloaded!

We can import the data frame using data(package = "package_name")

And assess the size of data frames using dim(data_frame) or preview the first few rows with head(data_frame)

Data frame format: Df[rows, columns]

```{r}
data(package = "palmerpenguins")

#assess the size of the penguins dataframe

dim(penguins)
# 344 represents the number of rows and 8 is the number of columns!

# look at the first few rows
head(penguins)
```

Let's try extracting columns from the data frame penguins
```{r}
# extract the second column and assign it to a variable

second_column <- penguins[2]

# extract the second column a different way

second_column_again <- penguins[, 2]

# extract the first row of the second column
single_cell <- penguins[1,2]

# extract column 3 and 10 by name 
name_col3 <- penguins["bill_length_mm"]
  
name_col10 <- penguins["year"]
```