---
title: "Tidyverse 7.0 worksheet"
author: "Katherine Brower (kbrower)"
date: "2024-06-11"
output: html_document
---
# 7.0 Tidyverse

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Tidyverse
Today we will start using the tidyverse package. Tidyverse is one of the most commonly used R packages and actually contains many packages!

Let's install and load the tidyverse package. Double check that for install.packages("") you have the "" marks and for library() there are no quotation marks. 

Remember that once you install a package once, you do not need to reinstall again! To use a package, you need to use library(package_name) at the top of your file.

```{r}
#install package
install.packages("tidyverse")
```

```{r}
library(tidyverse)
library(palmerpenguins)
```

Importing the penguins data from the palmerpenguins package
```{r}
data(package = "palmerpenguins")
```  

```{r}
head(penguins)
```
We can extract columns using select()
variable <- select(df, helper_function) or select(df, column names)
```{r}
# Let's select three columns using tidyverse: flipper_length_mm, body_mass_g, 
# and year

sel_columns <- select(penguins, flipper_length_mm, body_mass_g, year)

# Select a range of columns from bill_length_mm containing everything between 
# and including body_mass_g

sel_columns2 <- select(penguins, bill_length_mm : body_mass_g)
```

We can also use helper functions: 
- starts_with() : starts with a prefix
- ends_with() : ends with a prefix
- contains() : contains a literal string
- matches() : matches a regular expression
- num_range() : a numerical range
- one_of() : variables in character vector
- everything() : all variables

```{r}
# Use the ends_with() function to select columns containing mm
sel_columns3 <- select(penguins, ends_with("mm"))

# Use the contains() function to select columns containing length

sel_columns4 <- select(penguins, contains("length"))


# Use the starts_with() function to select columns containing 
sel_columns_5 <- select(penguins, contains("bill"))

# Can also combine many conditions together using | and & symbols!

sel_columns_lots <- select(penguins, contains("Species") | ends_with("mm"))
```

We can extract rows with the filter() function. This function works the same way as select() does for columns

```{r}
# filter rows for only male penguins
male_rows <- filter(penguins, sex == "male")
```

# Combining commands
Keep in mind that there are many ways to accomplish the same goal in R!

Intermediate objects can be very useful if you want to save or keep "intermediate" steps before the final object.
```{r}
# let's try choosing rows with filter using our sel_columns_lots dataframe as an intermediate object

# keep only "Adelie" penguin data
filtered_rows <- filter(sel_columns_lots, species == "Adelie")

# keep only penguins (of all types) with bill_length_mm over 20mm

filtered_rows2 <- filter(sel_columns_lots, bill_length_mm > 20)

# combine two conditions: keep only penguins with flipper length under 185 mm and Chinstrap

filtered_rows3 <- filter(sel_columns_lots, flipper_length_mm < 185 & species == "Chinstrap")
```

We can also combine them together as a nested function:
```{r}
nested_df <- select(filter(penguins, species == "Gentoo"), bill_length_mm, year, sex)
```