---
title: "Data Manipulation 9.0"
author: "Katherine Brower (kbrower)"
date: "2024-06-11"
output: html_document
---

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Split - Apply - Combine
We can also use the tidyverse package to do what we call a "split-apply-combine" strategy. This is really helpful because you can organize big datasets into more organized pieces and then put it back together.

For categorical data, it can be helpful to count members of a group.
We can do this by group_by() and tally()

```{r} 
# Count the number of penguins of each sex
penguins %>%
  group_by(sex) %>%
  tally()

# Currently we have 11 NA penguins included, we should remove these and 
# create a new table
clean_penguins <- penguins %>%
  group_by(sex) %>%
  filter(!is.na(sex)) %>%
  tally()
```

We can also use group_by() with summarize(). The main difference is that the non-grouping and non-output columns are removed

```{r}
# summarize the average bill_length_mm (in mm) by sex
bill_df <- penguins %>%
  group_by(sex) %>%
  filter(!is.na(sex)) %>%
  summarize(
  mean_bill_length =
      mean(bill_length_mm, na.rm = TRUE)
  )
```

# Saving data to a file
You can save results to a file using the write_csv() function. This can
be useful for later applications
```{r}
# save bill_df as a .csv file in our data directory
write_csv(bill_df, "data/bill_df.csv")
```

# count() & arrange()
The count() function can be very helpful since it combines group_by() and tally() together

The arrange() function sorts the data. To sort by ascending numbers: arrange(n) and if you want to sort by descending numbers: arrange(desc(n))

```{r}
# counting distinct number of penguin species in ascending order
penguin_counts <- penguins %>%
  count(species) %>%
  arrange(n)

penguin_counts

# counting distinct number of penguin species in descending order
penguin_counts_desc <- penguins %>%
  count(species) %>%
  arrange(desc(n))

penguin_counts_desc
```

# Using %in% operator
Sometimes we only want to include a specific value, the %in% operator is very helpful! This also works in a similar way to OR statements

```{r}
# extract data specifying which species to keep: Adelie and Chinstrap

penguins_reduced <- penguins %>%
  filter(
    species %in%
      c("Adelie", "Chinstrap")
  )

# this will create the same result:

penguins_reduced_same <- penguins %>%
  filter(species == "Adelie"|
         species == "Chinstrap")
```

## Trying these concepts with the new clinical dataframe
Let's try these functions with the clinical dataframe!
- group_by() and tally()
- group_by() with summarize()
- write_csv()
- count() and arrange()
- %in%

```{r}
# Count the number of each ajcc pathological stage and remove the NAs.
# ajcc pathological stage is used to determine to what the first course of 
# treatment will be for a patient

clinical %>%
  group_by(ajcc_pathologic_stage) %>%
  filter(!is.na(ajcc_pathologic_stage)) %>%
  tally()
```

Let's try with group_by() and summarize() together
```{r}
# summarize the average bill_length_mm (in mm) by sex
# summarize the average age at diagnosis (in days) by tissue or organ of origin
ERCC2_age_df <- clinical %>%
  group_by(tissue_or_organ_of_origin) %>%
  filter(!is.na(tissue_or_organ_of_origin)) %>%
  summarize(
  mean_age =
      mean(age_at_diagnosis, na.rm = TRUE)
  )

ERCC2_age_df

# It can be difficult to tell how old someone is by days. Let's add a new column
# to the ERCC2_age_df called mean_age_years and call the new dataframe 
# ERCC2_year_df. Remember that we can use the mutate() function to create columns.

ERCC2_year_df <- ERCC2_age_df %>%
  mutate(mean_year = mean_age / 365)
```

Let's save the new data frame that includes both mean age in days and years in
a new csv file called ERCC2_year_df.csv
```{r}
# save  as a .csv file in our data directory
write_csv(ERCC2_year_df, "data/ERCC2_year_df.csv")
```

count() and arrange()
```{r}
# counting distinct number of tissues of organs of origin in ascending order
tissue_organ_counts <- clinical %>%
  count(tissue_or_organ_of_origin) %>%
  arrange(n)

tissue_organ_counts
```

Using %in%
```{r}
# extract data specifying which year_of_diagnosis to keep: 2010 - 2013

clinical_reduced <- clinical %>%
  filter(
    year_of_diagnosis %in%
      c(2010:2013)
  )

clinical_reduced
# this will create the same result:

clinical_reduced_same <- clinical %>%
  filter(year_of_diagnosis == 2010|
         year_of_diagnosis == 2011 |
         year_of_diagnosis == 2012 |
         year_of_diagnosis == 2013)

clinical_reduced_same
```