---
title: "Piping 8.0 worksheet"
author: "Katherine Brower (kbrower)"
date: "2024-06-11"
output: html_document
---

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

Loading necessary packages
```{r}
library(tidyverse)
library(palmerpenguins)
```

Loading data from palmer penguins
```{r}
data(package = "palmerpenguins")
```

## Piping
Piping is part of the tidyverse package! This helps streamline and use the output of one command and put that result in the next command.

The piping operator is %>%

"." is very useful to pass the object calculated immediately before to the current function

```{r}
# use piping to create a df containing only female penguins (row) and select for columns year, bill_length_mm, and island

piped <- penguins %>%
  filter(., sex == "female") %>%
  select(., year, bill_length_mm, island)
  
# use piping to create a df containing only penguins with flipper_length_mm over 190mm and select for columns species, sex, and year

piped2 <- penguins %>%
  filter(., flipper_length_mm > 190) %>%
  select(., species, sex, year)
```

If we want to add a new column, we can use the function mutate()

```{r}
# use piping to create a new column called bill_length_cm where the bill_length_mm is converted into cm
bill_cm_df <- penguins %>%
  mutate(bill_length_cm = bill_length_mm * 0.1)


# use piping to create a new column called mass_lb where the mass_g is converted in lb

mass_lb_df <- penguins %>%
  mutate(mass_lb = body_mass_g * 0.002)
```

##Exploring the new dataframe called clinical

First we need to utilize the correct package (tidyverse) to read in csv files

```{r}
#load the tidyverse package
library(tidyverse)
```

Now we'll read in the dataframe from the clinical.csv file
```{r}
# importing in the new dataframe
clinical <- read_csv("data/clinical.csv")
```

Let's recall a few things we can do to become acuqainted with our dataset:
- determine the dimensions
- column names
- look at the first 6 rows of the dataframe
- determine the datatype within a column
- determine how many column names include the same word (ex. mutation)
- determine if there is any NA values within the data and how many in total

```{r}
#determine the dimensions (number of rows, number of columns)
dim(clinical)

# only number of rows
nrow(clinical)

# only number of columns
ncol(clinical)
```

```{r}
# look at the column names. Many of these column names may sound new to you!
# if you're interested in learning more, you can use google or the dataframe
# documentation. Many publicly available dataframes have 
colnames(clinical)
```

```{r}
# let's look at the first 6 rows of the data frame. This function is really
# helpful for becoming more acquainted with the contents of the data frame
head(clinical)
```

```{r}
# now that we have seen a general overview of the size and what content is
# inside, let's focus specifically on what data types we have

# to call for a column the format is: dataframe$column_name
# example let's look at year_of_birth
class(clinical$year_of_birth)

# Tobacco.smoking.history
class(clinical$Tobacco.smoking.history)
```

```{r}
# determine if there are any NA values within the data

# example column: sample_type
is.na(clinical$sample_type)
```

```{r}
# example column: pack_years_smoked
is.na(clinical$pack_years_smoked)

# we can also determine the total number of NA values within a column by using
# the sum function
sum(is.na(clinical$pack_years_smoked))
```

## Practice manipulating data in the clinical dataframe
The clincial dataframe is much larger than our previous penguin dataframe! We
can look into a lot more and combine different variables (columns) to see if
there are any patterns.

Let's make a new subset dataframe called smoke_clinical that includes 5 columns:
  - primary_diagnosis
  - Vital.status
  - Tobacco.smoking.history
  - pack_years_smoked
  - cigarettes_per_day
```{r}
smoke_clinical <- clinical %>%
  select(., primary_diagnosis, Vital.status, Tobacco.smoking.history,
         pack_years_smoked, cigarettes_per_day)
```

Using a bigger dataframe still utilizes the same functions and piping!