# 11.0 Intro to plots

We will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

```{r}
library(tidyverse)
library(palmerpenguins)
penguins <- penguins


```

### Recap: `filter`, `select`, `mutate`, `group_by`, `summarize` all with piping `%>%`

```{r}
adelie_body_mass_by_island <- penguins %>%
  filter(., species == "Adelie") %>%    #Filter only Adelie species
  select(., species, island, body_mass_g) %>%  #Select species, island, and body_mass_g column
  mutate(., body_mass_kg = body_mass_g / 1000 ) %>%  #Make new column: body_mass_kg
  group_by(island) %>%   # Group by island
  summarise(mean_body_mass_kg = mean(body_mass_kg, na.rm=TRUE)) #Summarise island group to calculate mean
  
adelie_body_mass_by_island
```

```{r}

```

## My first plot

We will use the `ggplot` package. The major components to a `ggplot` graph are:

-   Data (the information we want to show)

-   Aesthetics (the axes)

-   Geometry (the shapes, the plot type)

-   Themes (what everything looks like)

```{r}

```

We can add a preset theme to our plot in a similar grammar structure.

```{r}

```

Here is a list of themes readily available:

-   `theme_gray()` The signature ggplot2 theme with a grey background and white gridlines, designed to put the data forward yet make comparisons easy.

-   `theme_bw()` The classic dark-on-light ggplot2 theme. May work better for presentations displayed with a projector.

-   `theme_linedraw()` A theme with only black lines of various widths on white backgrounds, reminiscent of a line drawing. Serves a purpose similar to `theme_bw()`.

-   `theme_light()` A theme similar to `theme_linedraw()` but with light grey lines and axes, to direct more attention towards the data.

-   `theme_dark()` The dark cousin of `theme_light()`, with similar line sizes but a dark background. Useful to make thin coloured lines pop out.

-   `theme_minimal()` A minimalistic theme with no background annotations.

-   `theme_classic()` A classic-looking theme, with x and y axis lines and no gridlines.

-   `theme_void()` A completely empty theme.

Try a different theme here:

```{r}

```

## 2. Bar plots

Barplots plots categorical data and makes numerical comparisons, usually counts, between categories.

We will use the clinical dataset for more plotting examples.

```{r}
clinical <- read_csv("data/clinical.csv", show_col_types = FALSE)
```

Plot categorical data in a bar plot for the `mutation.in.TP53` data. For a bar plot, we use `geom_bar()` as the geometry type.

```{r}

```

## 3. Histogram

Histogram plots the range of a set of numerical values.

Plot a histrogram showing the range of `height`. For a histogram, we use `geom_histogram()` as the geometry type.

```{r}

```

Changing the number of bins: `bins=10`

```{r}

```

Changing the width of the bins: `binwidth=10`

```{r}

```

## 4. Scatter plots

Let's make a scatter plot to compare the number of cigarettes per day (`cigarettes_per_day` column) with the number of ERCC2 mutations (`count.ERCC2.signature.mutations` column.)

```{r}

```

We can alter the coordinate range to show a specific region using `coord_cartesian(xlim = c(0, 10), ylim = c(0, 400))`

```{r}

```

We can use pipes to remove NAs to clean the plot. Recall this from the piping lesson:

```{r}

```

This input can go directly into `data=`

```{r}

```

## 5. Box plots

Let's make a scatter plot to compare `Current smoker` and `Lifelong Non-smoker` (both in `Tobacco.smoking.history` column) with the number of ERCC2 mutations (`count.ERCC2.signature.mutations` column.)

First let's filter the data to just our data of interest:

```{r}
clinical <- read_csv("data/clinical.csv", show_col_types = FALSE)

```

Plot the filtered dataset:

```{r}

```

## Saving plots

After creating your plot, use `ggsave` to save your plot.

```         
ggsave("my_plot.png")
```

```{r}

```
