---
title: "Intro to Plots"
author: "Hanako and Matthew"
date: "8/13/2024"
output: html_document
---

# 11.1 Intro to Plot Exercises

### load library/package

```{r}
library(tidyverse)
```

### import data

```{r}
clinical <- read_csv("data/clinical.csv")
```

### Recall

The general formula for making plots is

```         
ggplot(data = dataframe) + aes(x = x_axis, y = y_axis) + geom_point() + theme()
```

## Part 1: Bar Plots!

Let's make a bar plot of mutations in `ggplot`, we want to see how many of the samples in `clinical` that have a mutation in RB1, that do not have a mutation in RB1, and how many are NA.

#### What is the x-axis

```{r}

```

#### What is the geometry (the plot type)

```{r}


```

#### Make the bar plot

```{r}


```

## Part 2: Histograms

Let's make a histogram of the distribution of `age_at_diagnosis` in the `clinical` dataset

#### What is the x-axis

```{r}

```

#### What is the geometry (the plot type)

```{r}


```

#### Make the histogram plot

```{r}


```

#### Make the histogram plot with 15 bins

```{r}

```

#### Actually, let's set the bin width to be 365 so we see everything in groups of 365 days

```{r}

```

## Part 3: Scatterplots

Let's make a scatter plot looking at the `number of cigarettes per day` and the `total number of single nucleotide variants`

#### Specify aesthetics: Let the x axis be `cigarettes_per_day` and the y axis be `total.number.Single.Nucleotide.Variants`

```{r}

```

#### What is the geometry type?

```{r}

```

#### Make the scatter plot

```{r}

```

#### Restrict the x axis from 0 to 10 and the y axis from 0 to 4500

```{r}

```

## Part 4: Box plots

Now we are going to make a box plot of ERCC2 mutations and compare them across current smokers and lifelong non-smokers.

### Refresher:Let's start with cleaning our data.

Use piping to: 

1. Subset the `clinical` data frame to include the columns: `barcode`, `Tobacco.smoking.history`, `total.number.Single.Nucleotide.Variants`. 
2. remove NAs in `total.number.Single.Nucleotide.Variants` 
3. filter rows to keep only `Current smoker` and `Lifelong Non-smoker` 
4. Call this `barcode_TSH_SNV_count_data`

```{r}

```

#### Specify aesthetics: What is the x-axis? What is the y-axis?

```{r}

```

#### What is the geometry type?

```{r}

```

#### Make the box plot

```{r}

  
```

#### Remove the outlier points.

```{r}




```

#### Add the jittered datapoints. Color the points a color of your choice.

```{r}




```

#### Restrict the y axis to 0 to 450

```{r}



```


