# 6.2 Dataframes Exercises

In this exericse, we will explore data about Starbucks coffee

As always, we will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

------------------------------------------------------------------------

## Load the dataset

Run the code chunk below. It will load download the data and assign it to the object `starbucks`. We will use this dataframe for the exercise.

```{r}
#### Do not alter the code below. This is a function that will download your dataset. 
starbucks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-12-21/starbucks.csv')
```

## Explore the dataframe

1.  Let's view the starbucks data frame you just downloaded using the `View()` function.

    ```{r}

    ```

2.  You can assess the size of the data frame using the function `dim(dataframe).`

    ```{r}

    ```

    **Answer: Number of rows = \_\_\_**

    **Answer: Number of columns = \_\_\_**

3.  You can preview the first few rows using the function `head(dataframe)`.

    ```{r}

    ```

4.  Use the `colnames()` function on the `starbucks` dataframe.

    ```{r}

    ```

5.  Use the `str()` function on the `starbucks` dataframe.

    ```{r}

    ```

    In your own words, what does the `str()` function do?

    ```         
    ```

6.  Finally, use the `summary()` function on the `starbucks` dataframe.

    ```{r}

    ```

    In your own words, what does the `summary()` function do?

    ```         
    ```

## Subsetting data

Now that we know how to make and look at dataframes, you can extract values from a dataframe using the `[]` accessor.

Recall the format is `dataframe[rows, cols]`.

1.  Extract the value at row 1, column 1 in `starbucks`

    ```{r}

    ```

2.  Try accessing the first row in `starbucks`

    ```{r}

    ```

3.  You can also use this to access an entire column by doing the opposite. Try accessing the first column in `starbucks`.

    ```{r}

    ```

4.  If you don't include the comma, R will assume you want the column. Try extracting the first column of `starbucks`, but this time don't use the comma.

    ```{r}

    ```

5.  Take a look at the output for `starbucks[,1]` and `starbucks[1]`, do they look the same?

    ```{r}

    ```

6.  You can also extract a range of rows and columns by using `:`. Try extracting the rows 1 through 3 in the second column of `starbucks`.

    ```{r}

    ```

7.  You can exclude things by using `-`. To exclude the first column, you can use `dataframe[-1]`. Try it below:

    ```{r}

    ```

8.  Now we can assess the data in the column by the name of its column, like this: `dataframe["column_name"]`. Subset the `product_name` from `starbucks`.

    ```{r}

    ```

9.  We can list all the column names we want to extract using commas `,` , like this: ``` dataframe["``column_name1``", "``column_name2``", "``column_name3``"]. ``` Subset the `calories`, `sugar_g`, and `caffeine_mg` from `starbucks`.

    ```{r}

    ```

## Accessing data

We can use the `$` the access column data into a vector. The general format is `dataframe$column_name`.

1.  Try accessing the `total_carbs_g` column data, using the `$` syntax.

    ```{r}

    ```

2.  Now access the column data in `cholesterol_mg` using the `$` syntax.

    ```{r}

    ```

3.  Assign the `sugar_g` column data to an object called `sugar`.

    ```{r}


    ```

4.  Now let's combine our knowledge of accessing data from dataframes with some functions we learned earlier.

    Calculate the range for `sugar`.

    ```{r}

    ```

5.  Now calculate the mean of `sugar`

    ```{r}

    ```

6.  What is the maximum sugar content?

    ```{r}

    ```

## Plotting data

1.  Assign the `calories` column data to an object called `drink_calories`.

    ```{r}

    ```

2.  Below is the code the making a scatter plot. There is a part in the code that specifies what data goes on the x-axis and what data goes on the y-axis. Put the `calories` and `sugar` vector in either `x=` or `y=`. Run the code chunk to make a scatter plot.

    ```{r}
    ggplot() + aes(x = , y = ) + geom_point()
    ```

3.  **Challenge**. Access another column from the `starbucks` dataframe. Assign that data into a vector object and update the plot code to compare the two columns.

    ```{r}

    ```

    ```{r}

    ```
