---
title: "My first Rmd file"
output: html_document
date: "2025-07-11"
---

# My first Rmd file

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: 

Today's date:

More notes:


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}

```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk. For any notes within a code chunk, don't forget to use the # symbol!

```{r}
print("Hello world")

```

We are going to use these Rmd files to write and save our code from now on!

## Rmd file practice/Intro to R Review

1. add 3 and 4 together, using the `+` operator.
    ```{r}


    ```

    Click on the run button to see the answer! It's the triangle button in the top right corner of the grey code chunk.

1. assign the value `15` to the object `weight_kg`
    ```{r}

    ```

    Once you run the code chunk, you should see `weight_kg` show up in the Environment window!

1. convert kilograms to pounds by multiplying `weight_kg` by 2.2
    ```{r}

    ```

1. assign a new object called weight_lb with the value you got above

    ```{r}


    ```

    Once you run the code chunk, you should see `weight_lb` pop up in the Environment window.


1. What is the data type of `weight_lb`?

    **Answer**: _________


5. re-assign `weight_kg` to 100
    ```{r}

    ```

    You should see the value for `weight_kg` change in your Environment window. 

1. round 3.14159 to 3 digits
    ```{r}


    ```


## Vector Review!
1. create a vector of your breakout group member's favorite animals called `animal_party`

    ```{r}


    ```


1. then add one animal to the front of the vector `animal_party`
    ```{r}


    ```

1. then add one marine animal to the end of the vector animal_party
    ```{r}
   

    ```

1. extract the 3rd animal and save it to an object called party_star
    ```{r}


    ```


1. unfortunately the marine animal wasn't able to make it to the party, remove them from the vector
    ```{r}


    ```



## Vector and Functions Review
1. Create a vector called `heights` in your vector, add the heights of your breakout room group members in inches.

    ```{r}

    ```

1. At the beginning of the vector, add your age.
    
    ```{r}

    ```


1. at the end of the vector, add the height of one of someone from your family

    ```{r}

    ```

1. what kind of data is in the vector heights?

    **Answer:** __________

1. Use 3 different functions with the vector heights (Ex. `max`, `min`, etc.) 

    Hint: Remember you can always review the slides or previous worksheets to see what functions we've used before! Or use the `?function_name` to try a new one

    ```{r}

    ```


## Excellent job!! 
Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  



