---
title: "Midpoint Review"
author: "Katherine Brower & Gabby Ferra"
date: "2023-08-10"
output: html_document
---
# Hello! This is our midpoint review R check in! Please fill out the
# code chunks and answer any questions with a comment using the # symbol.

# installing packages and loading packages into the rmd file
#      1) how do you install the tidyverse package? (This format is also how we install other packages besides tidyverse)
#      2) how do you load a package into the file?
```{r}
#make sure that the package name is in "" 
#1) install.packages("tidyverse")

# 2) package name is not in ""
library(tidyverse)
```

# import the data
# 3) import clinical.csv (from the pathway: "data/clinical.csv"make sure that you have 
# copied the data folder into the correct place to utilize the csv file!)
```{r}

# 3) your code here
clinical <- read.csv("data/clinical.csv")

```

# subsetting dataframes rows and columns
# 4) Filter clinical.csv to only have tumor samples
# 5) Subset clinical.csv to 4 columns of your choice

```{r}

# 4) 
filtered_rows <- filter(clinical, sample_type == "Primary Tumor")

# 5) your code here...
# way 1
clinical_subset <- clinical %>%
  select(project_id:sample_type)

# way 2
col4 <- select(clinical, barcode:primary_diagnosis)

# way 3
four_columns <- select(clinical, gender, height, year_of_diagnosis, sample_type)

# combine question 4 & 5 together
samples <- select(filter(clinical, sample_type == "Primary Tumor"), race, ethnicity, ajcc_pathologic_stage, height)

```
# creating vectors
# 6) Create a vector by setting it equal to a column in clinical.csv (choose a column made up of numbers, not words) hint: to see the columns, run View(clinical.csv)
```{r}

# 6) your code here...
#way 1
year_of_birth <- clinical$year_of_birth

#way 2
year <- select(clinical, year_of_birth)

#way 3
vector_try <- c(clinical$year_of_birth, clinical$height)

#the below code will not work to create the vector:
vector <- c(clinical, year_of_birth)
```
# determining data type and using various math functions (mean, median, minimum)
# 7) What is the data type in your vector?
# 8) With your vector from #6, find the mean median, mininimum, and maximum of the vector
```{r}

# 7) your code here...
class(vector_try)

# 8) your code here...
mean_value <- mean(year_of_birth, na.rm = TRUE)
median(year_of_birth, na.rm = TRUE)
min(year_of_birth, na.rm = TRUE)
max(year_of_birth, na.rm = TRUE)
```
# using na.rm and is.na
# 9) Are there any NAs in your vector? 
# 10) If so, remove them.
```{r}

# 9) Yes

# 10) your code here...
year_of_birth <- year_of_birth[!is.na(year_of_birth)]

```

# 
#Vector 1
# * create a vector of your breakout room group member's favorite animals called animal_party
# * then add one animal to the front of the vector animal_party
# * then add one marine animal to the end of the vector animal_party
# extract the 3rd animal and save it to an object called party_star
```{r}

# your code here...
animal_party <- c("tiger", "elephant", "pig", "bird", "manatee", "cat", "tiger")

animal_party <- c("dolphin", animal_party)

animal_party <- c(animal_party, "jellyfish")

party_star <- animal_party[3]

num_of_starfish <- length(animal_party[animal_party == "starfish"])
```

```{r}
num_of_tiger <- length(animal_party[animal_party == "tiger"])
num_of_tiger
```
# Additional fun with the animal party!
# What are 2 other things/functions you could do to add on or edit the vector animal party? Add them below!

```{r}
animal_party <- c(animal_party, "starfish")
animal_party <- animal_party[-6]
animal_party[6] <- NA

```
# Vector 2
# Create a vector called heights
# in your vector, add the heights of your breakout room group members
# at the beginning of the vector, add your age
# at the end of the vector, add the height of one of someone from your family

```{r}
heights <- c(5.3, 5.6, 5.3, 5.2, 5.6)
heights <- c(27, heights)
heights <- c(heights, 5)
heights
```


# Stats on Vector 2
# calculate the mean, median, maximum, and minimum of the vector heights
# what kind of data is in the vector heights?
# Add one more thing/function that you can do with the vector heights!
```{r}
mean(heights)
median(heights)
max(heights)
min(heights)
```

```{r}
class(heights)
length(heights)
heights <- heights[-4]
heights
```