---
title: "Intro to Plots: Bar plots and Histograms"
author: "Hanako Osuga"
date: "2/1/20224"
output: html_document
---

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# Bar plots practice
# the following are some categorical data in our clinical data set: tissue_or_organ_of_origin, sample_type, primary_diagnosis, ajcc_pathologic_state, gender

#lets practice making a bar plot looking at the count of records by primary_diagnosis
#call this plot diagnosis_plot
```{r}
diagnosis_plot <- ggplot(data= clinical,
                         mapping = aes(primary_diagnosis)) +
  geom_bar(stat="count")
diagnosis_plot
```

#make the same code with our alternative formatting
#call this diagnosis_2
```{r}
diagnosis_2 <- ggplot(clinical)+
    geom_bar(aes(x=primary_diagnosis), stat="count")
diagnosis_2
```


#pick another categorical variable from the list and make your own bar plot!
#what are you intersted in learning about our clinical data set?
```{r}
#this is an example with the column gender

gender_plot <- ggplot( data = clinical,
                       mapping = aes(x= gender))+
  geom_bar(stat="count")
gender_plot
```

#save this plot in the plots directory you created previously
#call this plot the name of the variable chose .pdf
```{r}

#this is an example with the column gender
ggsave(filename = "plots/gender.pdf",
       plot = gender_plot,
       width = 4, height = 3)

```


#Histogram practice!
#these are some numerical data columns in our clinical dataset: year_of_birth, age_at_diagnosis, year_of_diagnosis, pack_years_smoked, cigarettes_per_day


#lets practice making a histogram with the column cigarettes_per_day
#call this plot cigarettes_plot
```{r}
cigarettes_plot <- ggplot(data = clinical, 
                          mapping = aes(x= cigarettes_per_day))+
    geom_histogram()
cigarettes_plot
```

#change the number of bins or the bin widht on this plot to make it a little easier to read
```{r}
cigarettes_plot <- ggplot(data = clinical, 
                          mapping = aes(x= cigarettes_per_day))+
    geom_histogram(bins = 10)
cigarettes_plot
```


#choose another numerical variable to make your own histogram
#what are you interested in learning about the clinical dataset?
#make sure to utilize the bins/binwidth to make your plot easy to read!
```{r}
#this is an example with year_of_diagnosis

diagnosis_year <- ggplot(data = clinical, 
                         mapping = aes(x=year_of_diagnosis))+
  geom_histogram(binwidth = 5)
diagnosis_year

```

#save this plot in the plots directory you created previously
#call this plot the name of the variable chose .pdf
```{r}

#this is an example with the column year_of_diagnosis
ggsave(filename = "plots/year_of_diagnosis.pdf",
       plot = diagnosis_year,
       width = 4, height = 3)

```


