---
title: "Intro to Plots: Bar plots and Histograms Challenge Problems"
author: "Hanako Osuga"
date: "2/14/20224"
output: html_document
---


# load library/package
```{r}
#run if you haven't already today
#library(tidyverse)
```

# import data
```{r}
#run if you haven't already today
#clinical <- read_csv("data/clinical.csv")
```

# make a plot of the column vital_status
# you will need to choose the type of plot and build your plot accordingly
# hint: you will also need to do some data cleaning 
```{r}
# this is an example of a bar plot

vital_status_noNA <- clinical %>%
  select(Vital.status, barcode, year_of_birth) %>%
  filter(!is.na(Vital.status)) 

vital_status_plot <- ggplot(data= vital_status_noNA, 
                            mapping = aes(x= vital_status_noNA$Vital.status))+
  geom_bar(stat= "count")

vital_status_plot


```

# save this plot in the plots directory
# remember to name it something descriptive!
```{r}
ggsave(filename = "vital_status.pdf",
       plot = vital_status_plot,
       width = 4, height = 3)
```




# make the other type of plot (the one you didn't choose above)
# choose a column from the clinical dataset that works for the plot type 
# don't forget to remove NAs
```{r}
# this is an example of a histogram, we chose the year_at_birth

birth_year_noNA <- clinical %>%
  select(year_of_birth, barcode) %>%
  filter(!is.na(year_of_birth))


birth_year <- ggplot(data = birth_year_noNA,
                     mapping = aes(x= year_of_birth))+
        geom_histogram(binwidth=10) #looking at distribution by decade (10 years)

birth_year

```


# save this plot in your plots directory
```{r}
ggsave(filename = "birth_year.pdf",
       plot = birth_year,
       width = 4, height = 3)
```


