

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# bar plot: adding a third variable
# make a bar plot, showing the number of samples in each ajcc pathologic stage
# call this bar_plot2
```{r}

bar_plot2 <- ggplot(clinical, 
                   mapping = aes(ajcc_pathologic_stage)) +
  geom_bar(stat="count") 
bar_plot2

```
# let's add a third field, tissue_or_organ_of_origin 
# call this bar_plot3
```{r}
bar_plot3 <- ggplot(clinical,
                    mapping = aes(ajcc_pathologic_stage, fill = tissue_or_organ_of_origin))+
  geom_bar(stat = "count")
bar_plot3

```


# histograms: changing bins 
# by default R stacked the two bars
# make the same plot but bring the bars side by side
```{r}
bar_plot4 <- ggplot(clinical,
                    mapping = aes(ajcc_pathologic_stage, fill = tissue_or_organ_of_origin))+
  geom_bar(stat = "count",
           position = "dodge")
bar_plot4
```

# scatterplots: changing colors
# this is the same scatterplot as last time, plotting count.ERCC2.signature.mutations
# and the number of cigarettes per day

# change the color of the data points to pink 
```{r}
plot1 <- ggplot(data = clinical,
                mapping = aes(x = cigarettes_per_day, 
                              y = count.ERCC2.signature.mutations)) + 
  geom_point(color = "pink") # add a layer of geometry 
plot1

```

# scatterplots: adding trend lines
# now add a trend line
# make the trend line red
```{r}
plot1 <- ggplot(data = clinical,
                mapping = aes(x = cigarettes_per_day, 
                              y = count.ERCC2.signature.mutations)) + 
  geom_point(color = "pink") +
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 400))+
  geom_smooth(color= "red")
  
plot1
```


# box plot: changing colors 
# this is the same box plot as last time, it's plotting the count.ERCC2.mutations for current smokers and lifelong non-smokers
# change the color of the box plot to red and the fill orange
# change the jitter color to black 
```{r}
ggplot(data = barcode_TSH_ERCC2_count_data,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations)) + 
  geom_boxplot(outlier.shape = NA, color = "red", fill = "orange") + # remove points generated by geom_boxplot()
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "black")  # change jitter color 

```

```

