---
title: "Plots 2.0 Challenge Problems"
author: "sfhart & Katherine Brower"
date: "08/15/2023"
output: html_document
---


## Data: The Cancer Genome Atlas (TCGA) bladder cancer sequencing and patient
##       characteristics


## Graphing using ggplot2

## Instructions:
## Work with a partner or two!:
## - We've looked at a number of variables in our RB1_clinical dataset. Choose
##   two variables (columns) you think might have an interesting relationship.
## - Note: 
##   - it may be wise to avoid relating 2 categorical variables (may not be able to plot it!)
## - Form a hypothesis about these variables and discuss with your group which plot type would best help you evaluate it.
## - Generate the plot using the techniques we learning in ggplot.
## - Change the color/fill of at least one part of the plot and use
##    coord_cartesian to "zoom" in/out as necessary.
## - Add informative x- and y-axis labels.
## - Use ggsave() to save to plot to an appropriate name in the plots directory (folder)


```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```

# Run the following code from lesson 6 to make RB1_clinical
```{r}
clinical <- read_csv("data/clinical.csv")
RB1_clinical <- clinical %>%
  select(tissue_or_organ_of_origin:total.number.Indels, contains("RB1")) %>%
  filter(sample_type == "Primary Tumor") %>%
  mutate(across(c(focal.deletion.in.RB1,mutation.in.RB1), #find cols that contain "RB1"
                ~replace_na(., "NaN"))) %>%  #replace the NAs with strings
  mutate(any.RB1.mut = focal.deletion.in.RB1 == "Del" |
           focal.deletion.in.RB1 == "Loss" |
           mutation.in.RB1 == "yes"
  ) %>%
  mutate(RB1_expr = as.numeric(RB1_expr)) #make sure expression is numeric just in case
```


```{r}
# [Add code here]
```





# PART 2

All exercises use the clinical object we created in the readings:

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)

# Load the data
clinical <- read.csv("data/clinical.csv")
```

In the code used to create boxplots in the readings:
    Code jitter before the boxplot. What happens?
    Layer the code block using theme(), modifying the axes labels and the legend, before theme_bw(). What is different?
    Given your observations after doing the above, does the order of layers matter?
```{r}

```
    
In the boxplot code in the reading, use theme_gray() (or theme_grey()) instead of theme_bw(). What do you observe when theme_gray() is present vs when it is absent in the code block? What can you conclude about default theme settings in ggplot2?
```{r}

```

In the Scatterplots and Histograms lesson, we created a scatterplot relating smoking to the number of ERCC2 signature mutations. Use what we learned in Lesson 7b to create a more presentable plot.
    Set the colors of the points to a column with continuous (numerical) data of your choice. What happens? Is there a trend?
    Set the colors of the points to be by cancer stage category (relevant column is called ajcc_pathologic_stage). Is there a trend?
    Use faceting to create the same plot for each of the cancer stages
```{r}

```

Create a plot of your choice using the data in clinical. Use what you learned today to add appropriate customization.
```{r}

```


