---
title: "Plots 2.0 Challenge Problems KEY"
author: "Hanako Osuga"
date: "2.21.24"
output: html_document
---

As always, load the package if you haven't already this session
```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```

Load the data if you haven't already this session
```{r}
clinical <- read_csv("data/clinical.csv")
```


PART 1: Error Messages

The following code will throw error messages. Work with your group to understand what the error message is saying and find the bug in the code!

# error in geom_bar()
```{r}

bar_plot <- ggplot(data = clinical, 
                   aes(x = gender, 
                       y = sample_type)) +
  geom_bar() +
  theme_grey() + 
    theme(axis.text.y = element_text(color = "black", size = 10),
          axis.text.x = element_text(size = 8, color = "black"),
          axis.title.y = element_text(color = "black", size = 8),
          axis.title.x = element_text(color = "black", size = 8) )+
      labs(title = "Sample type vs Gender",
           x = "Gender",
           y = "Sample Type")
bar_plot

# ANSWER: this code plots too many variables for the plot type, bar plots generally
# show a count of a single variable, get rid of the y variable to get the code to run

```


# error: unexpected symbol
```{r}
ggplot(data = clinical),
       aes(x = year_of_birth)) +
  geom_bar()

# ANSWER: There is an extra ) after y=gender, remove it to get the code working!

```


# error: object not found
```{r}

diagnosis_year <- ggplot(data = clinical, 
                         mapping = aes(x=year_diagnosis))+
  geom_histogram(color = "#7fb366", binwidth = 5)
diagnosis_year

# ANSWER: there's a typo in year_diagnosis, it should be year_of_diagnosis

```


# error: cannot open file
```{r}

ggsave(filename = "plots/vital_status.pdf",
       plot = vital_status_plot,
       width = 4, height = 3)

# ANSWER: the directory named "plot" doesn't exist, change it to the actual 
# name of your folder

```



PART 2: Odd Plots
The following code won't throw error messages but the plots look a little odd. Work with your group to fine tune these plots and make them more understandable!

# missing data? 
# this one won't throw an error but the plot looks... lacking
```{r}

#cleaning the dataset 
non_smoker_clinical <- clinical %>%
  filter(Tobacco.smoking.history== "ND" | Tobacco.smoking.history == NA)

# making the plot
plot_scatter <- ggplot(data = non_smoker_clinical, 
                       aes(x = cigarettes_per_day,
                            y = pack_years_smoked)) + 
  geom_point(color = "#8ad6cc") +
  theme_grey()+
    theme(axis.text.y = element_text(color = "black", size = 10),
          axis.text.x = element_text(size = 8, color = "black"),
          axis.title.y = element_text(color = "black", size = 8),
          axis.title.x = element_text(color = "black", size = 8) )+
      labs(title = "Cigarettes_per_day vs pack_years_smoked",
           x = "cigarettes per day",
           y = "pack years smoked")
plot_scatter


#ANSWER: the data set contains non-smokers, so they wouldn't have any data on cigarettes 
# smoked, to get data add ! in the filter step to get the data on smokers
    
```


# planning your plot
# is this plot the best plot to showcase this data?
```{r}

ggplot(data = clinical, aes(x= ajcc_pathologic_stage,
                            y = gender)) +
  geom_point(color = "#c06df8" )+ 
  theme_grey()+
  theme(axis.text.y = element_text(color = "black", size = 10),
        axis.text.x = element_text(size = 8, color = "black"),
        axis.title.y = element_text(color = "black", size = 8),
        axis.title.x = element_text(color = "black", size = 8) )+
  labs(title = "gender and cancer stage")

# ANSWER: you are trying to plot two categorical variables so a scatter plot may not 
# be the best plot for this. What would work better?

```


# where are my labels?
# this code won't throw an error message, but it doesn't include the axis labels you coded for, what's going on? 
```{r}
my_plot <- ggplot( data = clinical, 
                   aes(x= year_of_birth, 
                       y= total.number.Indels)) +
  geom_point(aes(color = ajcc_pathologic_stage) ) +
  coord_cartesian(ylim = c(0,30)) +
  theme_grey()
    theme(axis.text.y = element_text(color = "black", size = 10),
        axis.text.x = element_text(size = 8, color = "black"),
        axis.title.y = element_text(color = "black", size = 8),
        axis.title.x = element_text(color = "black", size = 8) )+
    labs(title = "Total Number of Indels by Year of Birth",
         x = "Year of Birth",
         y = "Total Number of Indels")
           
my_plot

# ANSWER: missing a + after theme_grey(), if you don't add the + R won't
# know that there is more information coming

```

