---
title: 'Lesson 6: Coding Exercises'
author: "Sam Hart & Katherine Brower"
date: "2023-08-10"
output: html_document
---

## Data: The Cancer Genome Atlas (TCGA) bladder cancer sequencing and patient
##       characteristics



## Part I: Counting and summarizing groups

## Instructions:
##  We are interested in how RB1--which is best known for contributing to eye
##  cancers--might contribute to bladder cancers. Let's continue with our expression
##  comparisons from last time!
##   - Group by "any.RB1.mut" and then determine the median, min, and max
##     of RB1_expr for each group using summarize.
##   - Group by cancer stage and then determine the median age of diagnosis *in years*
##     Hint: Remember to use View() or names() if you're not sure of the column
##      names to use here. Remember you can nest mathematical functions just like
##      R functions or the two together. :)


```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```

## Load the data

```{r}
clinical <- read_csv("data/clinical.csv")
```

## Continue with RB1-relevant subset of data

```{r}
RB1_clinical <- clinical %>%
  select(tissue_or_organ_of_origin:total.number.Indels, contains("RB1")) %>%
  filter(sample_type == "Primary Tumor") %>%
  mutate(across(c(focal.deletion.in.RB1,mutation.in.RB1), #find cols that contain "RB1"
                ~replace_na(., "NaN"))) %>%  #replace the NAs with strings
  mutate(any.RB1.mut = focal.deletion.in.RB1 == "Del" |
           focal.deletion.in.RB1 == "Loss" |
           mutation.in.RB1 == "yes"
  ) %>%
  mutate(RB1_expr = as.numeric(RB1_expr)) #make sure expression is numeric just in case
```

##   - Group by "any.RB1.mut" and then determine the median, min, and max
##     of RB1_expr for each group using summarize.

```{r}
RB1_stats <- RB1_clinical %>%
  group_by(any.RB1.mut) %>%
  summarize(
    median.RB1_expr = median(RB1_expr, na.rm = TRUE),
    min.RB1_expr = min(RB1_expr, na.rm = TRUE),
    max.RB1_expr = max(RB1_expr, na.rm = TRUE)
  )

# # A tibble: 2 × 4
#    any.RB1.mut median.RB1_expr min.RB1_expr max.RB1_expr
#    <lgl>                 <dbl>        <dbl>        <dbl>
#  1 FALSE                  7.83        0.696         41.7
#  2 TRUE                   3.19        0.385         19.0

## To find col name:

names(RB1_clinical) #OR
View(RB1_clinical)
```

##   - Group by cancer stage and then determine the median age of diagnosis *in years*
##     Hint: Remember to use View() or names() if you're not sure of the column
##      names to use here. Remember you can nest mathematical functions just like
##      R functions or the two together. :)

```{r}
RB1_median_age <- RB1_clinical %>%
  group_by(ajcc_pathologic_stage) %>%
  summarize(
    median.age_diagnosis_yrs = median(age_at_diagnosis / 365)
  )

## A tibble: 5 × 2
#   ajcc_pathologic_stage median.age_diagnosis_yrs
#   <chr>                                    <dbl>
# 1 Stage I                                   73.5
# 2 Stage II                                  67.3
# 3 Stage III                                 69.7
# 4 Stage IV                                  69.0
# 5 NA                                        NA

```


## Part II: Saving data and more counting options

## Instructions:
##  - Save RB1_clinical to a new csv called "RB1.csv" in the directory data
##     Hint: If you're not sure what path to enter, use the read_csv statement
##      above as a template.
##  - Group by the column "any.RB1.mut" and count the number of patients with each cancer stage using tally()
##  - Filter race for Asian, Black, or White patients using the %in% operator;
##    then count the number of patients with each cancer stage using count()


##  - Save RB1_clinical to a new csv called "RB1.csv" in the directory data
##     Hint: If you're not sure what path to enter, use the read_csv statement
##      above as a template.
```{r}
RB1_clinical %>%
  write_csv(., file = "data/RB1.csv")
```


##  - Group by the column "any.RB1.mut" and count the number of patients with each cancer stage using tally()
```{r}
RB1_num_patients <- RB1_clinical %>%
  group_by(any.RB1.mut, ajcc_pathologic_stage) %>%
  tally()

## This one can be tricky since you have to use both in the group_by, which might
## not be clear from how it's written.

#    any.RB1.mut ajcc_pathologic_stage     n
#    <lgl>       <chr>                 <int>
#  1 FALSE       Stage I                   4
#  2 FALSE       Stage II                102
#  3 FALSE       Stage III               125
#  4 FALSE       Stage IV                118
#  5 FALSE       NA                        1
#  6 TRUE        Stage II                 28
#  7 TRUE        Stage III                17
#  8 TRUE        Stage IV                 18
#  9 TRUE        NA                        1

```

## Determine what these race categories are coded as; it's easier to copy and
## paste anyway :)
```{r}
unique(RB1_clinical$race)
# [1] "white"                     "asian"                     "not reported"
# [4] "black or african american"

```

##  - Filter race for Asian, Black, or White patients using the %in% operator;
##    then count the number of patients with each cancer stage using count()
```{r}
RB1_by_race_count <- RB1_clinical %>%
  filter(race %in% c("white","asian","black or african american")) %>%
  count(ajcc_pathologic_stage)

# ajcc_pathologic_stage      n
# <chr>                   <int>
# 1 Stage I                   4
# 2 Stage II                125
# 3 Stage III               135
# 4 Stage IV                132
# 5 NA                        1

```


## What would you add to the above code if you wanted to look at stage distribtions
## for the three race options?
```{r}
RB1_by_stage_distribution <- RB1_clinical %>%
  filter(race %in% c("white","asian","black or african american")) %>%
  group_by(ajcc_pathologic_stage) %>%
  count(ajcc_pathologic_stage)
```


