---
title: "Intro to Plots: Bar plots and Histograms"
author: "Hanako"
date: "2/1/20224"
output: html_document
---

# Intro to Plotting: Bar plot and histograms

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# Part 1: Bar Plots!
# let's make a bar plot of mutations.in.ggplot, we want to see how many of the samples in clinical
# have a mutation in TP53, do not have a mutation in TP53, and how many are NA

# basic ggplot step 1: bind data to the plot
# call your plot, bar_plot
```{r}

```

# basic ggplot step 2: specify aesthetics (axes)
# the x axis is mutation.in.TP53
```{r}


```

# basic ggplot step 3: add a layer of geometry
```{r}


```

# write the above command in the alternative formatting
```{r}


```



# create a directory to store our plots
```{r}

```

#save your plot!
```{r}

```




# Part 2: Histograms
# let's make a histogram of the distrbution of height in the clinical dataset

# basic ggplot step 1: bind the data to the plot
# name the plot histogram
```{r}


```

# basic ggplot step 2: specfiy the aesthetics (axes)
```{r}

```

# basic ggplot step 3: add a layer of geometry
# we want to make this with the default 30 bins
```{r}

```


# on second thought, let's make it with half the number of bins (15)
```{r}

```

# actually, let's set the bin width to be 10 so we see everything in groups of 10 inches
```{r}

```

# write the above command using the alternative formatting
```{r}


```

#save your plot in the plots directory you created earlier
```{r}

```


