---
title: "Intro to Plots: Bar plots and Histograms"
author: "Hanako Osuga"
date: "2/1/20224"
output: html_document
---

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# Bar plots practice
# the following are some categorical data in our clinical data set: tissue_or_organ_of_origin, sample_type, primary_diagnosis, ajcc_pathologic_state, gender

#lets practice making a bar plot looking at the count of records by primary_diagnosis
#call this plot diagnosis_plot
```{r}

```

#make the same code with our alternative formatting
#call this diagnosis_2
```{r}

```


#pick another categorical variable from the list and make your own bar plot!
#what are you intersted in learning about our clinical data set?
```{r}



```

#save this plot in the plots directory you created previously 
#call this plot the name of the variable chose .pdf
```{r}



```


#Histogram practice!
#these are some numerical data columns in our clinical dataset: year_of_birth, age_at_diagnosis, year_of_diagnosis, pack_years_smoked, cigarettes_per_day


#lets practice making a histogram with the column cigarettes_per_day
#call this plot cigarettes_plot
```{r}



```

#change the number of bins or the bin widht on this plot to make it a little easier to read
```{r}



```


# now lets make a subset of the clinical data to make the same plot but for 
# Current Smokers that are alive 

# combine what we learned about piping and data manipulation to create a new
# subset called clincal_smokers and use that data frame to make the plot
```{r}



```




#choose another numerical variable to make your own histogram
#what are you interested in learning about the clinical dataset?
#make sure to utilize the bins/binwidth to make your plot easy to read!
```{r}




```


# make the same plot but with a new subset of data that only contains data
# on individuals who was born after 1950 and select only the relevant
# columns for your plot
```{r}



```



#save this plot in the plots directory you created previously 
#call this plot the name of the variable chose .pdf
```{r}




```


