---
title: "11.0 Scatterplots and Box Plots"
author: "Sam Hart & Katherine Brower"
date: "7/17/2022"
output: html_document
---

# Introduction to R: Scatterplots and Box Plots

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```


# Part 1: Scatterplots
# let's make a scatter plot looking at the number of cigarettes per day and the count of
# ERCC2 mutations

# basic ggplot step 1: bind data to the plot
# call this plot1
```{r}

```

# basic ggplot step 2: specify aesthetics
# the x axis is cigarettes_per_day and the y axis is count.ERCC2.signature.mutations
```{r}

```

# basic ggplot step 3: add a layer of geometry
```{r}

```

# basic ggplot step 3: alternative way to add a layer of geometry
```{r}

```

# basic ggplot step 4: alter the coordinate range
# restrist the x axis from 0 to 10 and the y axis from 0 to 100 
```{r}

```

# write the command in the alternative formatting for ggplot 
```{r}



```


# we need to cleano out data! 
# create a subsetted data frame of the columns: barcode, cigarettes_per_day, count.ERCC2.signature.mutations and remove NAs
# call this new subset barcode_CPD_ERCC_count_data 
```{r}



```

#now that our data is clean,
# create a scatterplot using the subsetted data frame from above (steps 1, 2 and 3)
```{r}



```

# let's remove the outliers to make the data easier to read
# zoom into a smaller x-axis range (exclude outliers) (step 4)
# limit the x axis to 0 to 10 and the y axis to 0 to 400
```{r}




```
          
          

# Part 2: Box plots
# now we are going to make a box plot of ERCC2 mutations and comparte them across
# current smokers and lifelong non-smokers

# let's start with cleaning our data
# subset another data frame to include the columns: barcode, Tobacco.smoking.history, count.ERCC2.signature.mutations
# remove NAs in ERCC2 signature mutations
# restrict to current smokers and lifelong non-smokers
# call this barcode_TSH_ERCC2_count_data
```{r}




```

# now that we have clean data, let's make our box plot
# create a boxplot from the subsetted dataframe above (steps 1, 2, and 3)
```{r}



```


# create boxplot, adjust y-axis range (step 4)
# restrict the y axis to 0 to 125
```{r}




```

# let's add the data points onto the box plot
# add  data, jittered (step 5)
```{r}




```

# there are a few outliers that make it hard to see the trends
# let's remove the outlier points from geom_boxplot()
# this time make the jitter data points purple
```{r}



```

