---
title: "Scatterplots and Boxplots Exercises"
author: "sfhart & Katherine Brower"
date: "08/14/2023"
output: html_document
---

## Data: The Cancer Genome Atlas (TCGA) bladder cancer sequencing and patient
##       characteristics
```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```


## Graphing using ggplot2

#  Instructions:
# Let's revisit the relationship between RB1 expression and the types of
# mutations in RB1 that we've seen in our bladder cancer samples!


# to start, let's remake the subset RB1_clinical that we made in the piping exercise
# RB1_clinical should have data from clinical that selects for the columns tissue_or_organ_of_origin through
# total.number.Indels that contain "RB1" (we want to find columns that contain RB1)
# it should also filter for any rows that has Primary Tumor as the sample type

```{r}




```

# let's add on to the code above to do the following:
# 1) replace the NAs in focal.deletion.in.RBA and mutation.in.RB1 with "NaN"
# 2) create a new column called "any.RB1.mut"; set it equal to having "Del" or "Loss" in focal.deletion.in.RB1 OR "yes" in mutations.in.RB1 
# 3) double check RB1_expr is a numeric, just in case

```{r}



```


# Now that we have the data we need, lets make our plot
#  Make a boxplot comparing RB1 expression between cancers that have any
#  RB1 mutation and those that do not. Use the RB1_clinical dataset you created previously.
#  Hint: Which column holds logical values for our RB1 mutations of interest? (e.g. use names() or View())
# Call this any_RB1_mut

```{r}



```

# Reflection Questions
##    We were aware of the difference in mean/medians, but what can we now
##      say about the range, outliers, etc.? What new questions/hypothesis can
##      we generate from these data?


##  - Then use coord_cartesian() to "zoom" to between 0 and 20 FPKM (the units
##    for this analysis)
```{r}




```
# don't forget to save your plot!
```{r}



```


#  Reflection Questions
#  How does this new view alter what questions we might generate?


# now lets use RB1_clinical to make a scatterplor
# make a scatterplot of RB1_expr and age_at_diagnosis
```{r}


```


# lets remove the outlier by zooming into 0 to 30
```{r}



```

# we still have that second outlier, change the coord_cartesian again to remove it
```{r}



```


# save the plot!
```{r}



```

