---
title: "11.0 data manipulation group worksheet"
output: html_document
date: "2024-08-13"
---

#### Setup: loading in packages and reading in the clinical dataset

```{r}
#Load in the necessary packages: tidyverse
library(tidyverse)
```

```{r}
# read in the clinical.csv file as an object called clinical
clinical <- read.csv("clinical.csv")
```


#### Piped dataframe practice and mutate()
```{r}
# create a piped dataframe to answer the question:
# What is the relationship between the height of patients having a mutation in TP53 or a mutation in EP300?

# rows and columns needed:
# row: 
# column:


# piped dataframe: call this object height_TP53mut_EP300mut_df


# mutate() question: In height_TP53mut_EP300mut_df we currently have the height data in cm, let's convert this to inches and call the new column height_cm. Name the new dataframe as heightCM_2013_TP53mut_EP300mut 

```

```{r}
# create a piped dataframe to answer the question:
# Is there a correlation between the age at diagnosis and E2F3 expression and SPTAN1 expression?

# rows and columns needed:
# row:
# column: 


# piped dataframe: call this object ageDiagnosis_E2F3expr_SPTAN1expr


# Currently, the age of diagnosis data is in days, let’s convert this into years and call the new column age_of_diagnosis_years


```

```{r}
# Is there a correlation between gender and ATM expression and FGFR3 expression for patients born in 2013?
# rows and columns needed:
# row:
# column: 

# piped dataframe: call this object gender_ATMexpr_FGFR3expr_2013

gender_ATMexpr_FGFR3expr_2013 <- clinical %>%
  filter(year_of_birth == "2013") %>%
  select(ATM_expr, FGFR3_expr, year_of_birth, gender)

# We’re interested in combining ATM expression and KDM6A expression together, let’s add these together and call the new column combo_expr
```

#### Using the clinical dataframe, determine how many of patients are classified in each ajcc_pathologic_stage using group_by() & tally()

```{r}

  
# which ajcc pathologic stage has the most patients?


```

#### Using the clinical dataframe, how many patients are classified in each tobacco smoking history category? Use group_by() and tally()
```{r}
 
```

#### Using the clinical dataframe, summarize the average year of birth by gender. call this avg_year_byGender

```{r}



# Our average year of birth is currently a numeric, let's round to the nearest whole number and put the result in a new column called yob_avg_rounded. Name the dataframe as avg_year_byGender_rounded_df


```

#### We want to save our new avg_year_byGender_rounded_df as a csv file. Call the new file, clinical_YOB_byGender.csv

```{r}

# the new file should be visible in your files tab on the bottom right panel
```

#### Using the clinical dataframe, summarize the average total number indels by race. Call this avg_total_indels_byRace

```{r}

```

#### Using the clinical dataframe, summarize the median count ERCC2 signature mutations by mutation in ERCC2. 
```{r}


```

#### We want to save our new ERC22_median_df as a csv file. Call the new file, clinical_ERCC2_median.csv
```{r}

```

#### Using the clinical dataframe, count the number of distinct races and arrange this in a descending order

```{r}

```


#### Using the clinical dataframe, count the number of distinct tissue_or_organ_of_origin and arrange in ascending order
```{r}


```

#### Using the %in% operator, make a new subsetted dataframe called clinical_reduced1 and filter for Tobacco smoking history and only include Lifelong Non-smoker and Current smoker

```{r}

```

#### Using the %in% operator, make a new subsetted dataframe called clinical_reduced2 and filter for ajcc_pathologic_stage and only include Stage II and Stage III
```{r}

```