---
title: "12.0 Plots 2.0"
author: "Hanako osuga"
date: "2/9/2024"
output: html_document
---

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# bar plot: adding a third variable
# make a bar plot, showing the number of samples in each ajcc pathologic stage
# call this bar_plot2
```{r}



```
# let's add a third field, tissue_or_organ_of_origin 
# call this bar_plot3
```{r}


```


# histograms: changing bins 
# by default R stacked the two bars
# make the same plot but bring the bars side by side
```{r}


```

# scatterplots: changing colors
# this is the same scatterplot as last time, plotting count.ERCC2.signature.mutations
# and the number of cigarettes per day

# change the color of the data points to pink 
```{r}



```

# scatterplots: adding trend lines
# now add a trend line
# make the trend line red
```{r}


```


# box plot: changing colors 
# this is the same box plot as last time, it's plotting the count.ERCC2.mutations for current smokers and lifelong non-smokers
# change the color of the box plot to red and the fill orange
# change the jitter color to black 
```{r}



```
