---
title: "Plots 2.0 Exercises"
author: "Hanako Osuga"
date: "2/24/2024"
output: html_document
---

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# boxplot from lesson 7a
```{r}
d_plot <- clinical %>% 
  select(barcode, Tobacco.smoking.history, count.ERCC2.signature.mutations) %>% 
  filter(!is.na(count.ERCC2.signature.mutations)) %>% 
  filter(Tobacco.smoking.history %in% c("Current smoker", "Lifelong Non-smoker"))

L7A_boxplot <- ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "blue")

L7A_boxplot
```


# set boxplot color = "red"
```{r}



```

# set boxplot fill = "red"
```{r}




```

# set boxplot fill to be colored by Tobacco smoking history
```{r}



```

# specify boxplot fill colors
```{r}



```

# specify boxplot fill and jitter points colors
```{r}



```

# use hex color codes
# choose your own or use "#fdae61" and "#abd9e9" 
```{r}



```

# add transparency to jittered points
```{r}



```

# save plot to an object p
# don't forget to add a title and axis labels
```{r}


```

# modify plot area and gridlines, use a new theme!
```{r}



```
  
# modify plot text features with theme() to make it more readable!
# try color = black, and size = 7
# you can put the axis at an angle if that makes it easier to read
```{r}



```

# remove legend
```{r}



```

# fix axis titles
```{r}



```

# create final_plot object
```{r}




```
 
# save plot to plot directory
# name it something descriptive!
```{r}
 
```


