---
title: "2.1_Data_Type_Exercises"
output: html_document
date: "2024-01-19"
---

welcome to the Data Types Exercises!

dont forget to take notes and record your commands from swirl here so you can save
your work to reference later!

which of these is not a data type in R (numerics, logical/boolean, character, irrational, integer)

which data type is TRUE?

what is the data type that is enclosed in ""?

what is the assignment operator?

now let's practice assigning and re-assigning objects in R

#create an object called agge that contains 16
```{r}

```

#whoops we made a typo, reassign agge to a new object called age
```{r}

```

as you assign and re-assign objects check your Environment pane to make sure these changes go through

#calculate age in days by multiplying it by 365 (lets ignore leap years)
```{r}

```


not we are going to do some more practice assigning values to objects and using them

#create an object called mass, assign it 47.5
```{r}

```

#create an object called width, assign it 122
```{r}

```

when assigning and re-assigning, you can reference already existing objects

#re-assign mass with a new value of mass* 2.0
```{r}

```

#re-assign width with the new value width-20
```{r}

```

#create a new object called mass_index, assign it the value mass/width
```{r}

```

what is the data type stored in mass_index?























