---
title: "2.2_Data_Types_Challenge_Problems"
output: html_document
date: "2024-01-19"
---

welcome to the challenge problems for the data types lesson!

in this problem set, you work at a bodega that sells candy, a popular item here is the chocolate bar

#assign an obeject called item with the value "chocolate bar"
```{r}

```

each chocolate bar costs $1.25

#assign the object price with the price of the chocolate bar as a numeric
```{r}

```

WA State is a state with sales tax of 6.5%

#calculate the total price of 1 chocolate bar with sales tax, assign thsi to with_tax
```{r}

```

#how much would it cost to buy 6 chocolate bars BEFORE tax?
```{r}

```

#how much would it cost to buy 6 chocolate bars AFTER tax?
```{r}

```

the price of the chocolate bar goes down to $1.15

#re-assign price to this new number
```{r}

```

#what is the new price with tax per chocolate bar, re-assign with_tax with this new value
```{r}

```

you receive a large order of chocolate bars, the quantity requested is seventy chocolate bars

#make a new object called quantitiy, and assign it 'seventy'
```{r}

```

you are asked to calculate the price of this large order
since you were storing these values as you went, you should be bale to just ask R to do price*quantity

#try this
```{r}

```

why will this get you an error?

what is the data type of price?

what is the data type of quantity?

#re-assign quantity to the correct data type to do the calculation
```{r}

```

#calculate the price after tax for this large order
```{r}

```


great work!





