---
title: "3.0_Functions"
output: html_document
date: "2024-01-20"
---

functions are built in complex tasks or a predefined set of code that allows you to repeat particular actions

an example of a function is round(), it rounds the number entered in the ().

in this case, the function is round(), and the number entered in the parentheses is the argument.

thinking back to the previous lesson, what date type has decimal points?



#try rounding the numeric 3.14 usign the function above by entering 3.14 as an argument
```{r}


```

you can specify the number of digits you want to round to by adding ", digits = number" after the first argument

#now try rounding 3.14159 to 2 digits
```{r}

```

if you ever forget what a function does you can eenter the function name after a ?
```{r}
?round
```

a help window will pop open in the bottom right corner of your Posit Cloud with infromation on the function

#using the round function with complete argument labels, enter x=3.14 and digits=0 as arguments into the function round()
```{r}

```

did you get the same answer as round(3.14)?

let's try switching order of the arguments

#copy your code from the last comand and switch the position of the arguments
```{r}

```

what happened? did you get the same answer?


switching the order doesn't change when you use argument labels (ex: x= and digits = )

#now let's try removing the argument labels (x= and digits= ) from your last line of code
```{r}

```


what happened? was your answer the same with the argument labels?

when you have argument labels the computer can understand what you are asking, but without the argument labels it will assume the default order or arguments

so argument order doesn't matter as much when you use labels, but if you don't have to be mindful of argument order (the order the arguments are entered)




