---
title: "8.0_piping.Rmd"
output: html_document
date: "2024-01-23"
---


# load library/package 

```{r}
# run me
library(tidyverse)
```

# import data (same as yesterday!)

```{r}
# run me
clinical <- read.csv("data/clinical.csv") 
```

# earlier we used nested commands to execute multiple commands at once 

```{r}
# select patients with Stage IV cancer, then select only the race, ethnicity, and stage columns
race_stage <- select(filter(clinical, ajcc_pathologic_stage == 'Stage IV'), race, ethnicity, ajcc_pathologic_stage)

```
Take a look at the code above, what is the function filter doing?



What is the inner most nest (layer) of code?


#lets do the same task as above, but with pipes

# first, start off the pipe with the pipe operator, %>%, call this piped
# then write the first layer filtering for ajcc_pathologic_stage == "Stage IV",  use. to reference the clinical dataset
# then write the layer for selecting the race, ethnicity, and stage columns, use > to reference the clinical dataset you filtered above
```{r}

```
If you want to read more about piping,  https://magrittr.tidyverse.org/



# extract race, ethnicity, and age at diagnosis from cases born prior to 1930, filter before you select your columns, call this piped2
```{r}

```


# do the same as above, but this time select your columms before filtering rows, call this piped3
```{r}

```

Great work!
