---
title: "8.0_piping.Rmd"
output: html_document
date: "2024-01-23"
---
# load library/package 

```{r}
# run me
library(tidyverse)
```

# import clinical dataset

```{r}
# run me
clinical <- read.csv("clinical.csv") 
```

# earlier we used nested commands to execute multiple commands at once 

```{r}
# select patients with Stage IV cancer, then select only the race, ethnicity, and stage columns
race_stage <- select(filter(clinical, ajcc_pathologic_stage == 'Stage IV'), race, ethnicity, ajcc_pathologic_stage)

```
Take a look at the code above, what is the function filter doing?



What is the inner most nest (layer) of code?


# let's do the same task as above, but with pipes

# first, start off the pipe with the pipe operator, %>%, call this piped
# then write the first layer filtering for ajcc_pathologic_stage == "Stage IV",  use. to reference the clinical dataset
# then write the layer for selecting the race, ethnicity, and stage columns, use > to reference the clinical dataset you filtered above
```{r}

```
If you want to read more about piping,  https://magrittr.tidyverse.org/

# extract race, ethnicity, and age at diagnosis from cases born prior to 1930, filter before you select your columns, call this piped2
```{r}

```

# do the same as above, but this time select your columms before filtering rows, call this piped3
```{r}

```

# what is the difference between piped2 and piped3? Write the answer as a comment!
```{r}

```

## create a subsetted dataframe containing the column range tissue_or_organ_of_origin through total.number.Indels and contains RB1
## call this subsetted dataframe RB1_clinical

```{r}
RB1_clinical <- select(clinical,
                       tissue_or_organ_of_origin:total.number.Indels,
                       contains("RB1"))
```

# filter RB1_clinical for just the "Primary Tumor" samples types in "sample_type"
```{r}


```


# filter RB1_clinical for patients who have either "Del" or "Loss" in focal.deletion.in.RB1 OR "yes" in mutations.in.RB1 
# hint: we can use the "|" operator for OR. 
```{r}

```

# use the pipe operator to combine the selecting and filtering steps above
  
```{r}

```

# Let's further explore the clinical dataframe with variables that you're intersted in!
# Make a subsetted dataframe called my_experiment containing at least 3 different columns of your choice.
# Use piping and filter for at least one condition! Describe the relationship or what you're looking at.
# Example: I'm interested in seeing if patients in this dataset that are taller than 170 have more
# expression in PPARG or ASXL2?

# Hints: Check if the columns you're choosing contain NA values! Also see how the data was recorded using
# the unique() function.
```{r}
# I'm interested in looking at: 
# columns:
# rows:

# piped dataframe:

```

# That was an interesting result! Let's try doing that again with a different combination of columns.
# This time, use a comparison operator (either & or |) 
```{r}
# I'm interested in looking at: 
# columns:
# rows:

# piped dataframe:
```
Great work!
