---
title: "8.0_piping.Rmd"
output: html_document
date: "2024-01-23"
---


# load library/package 

```{r}
# run me
library(tidyverse)
```

# import data (same as yesterday!)

```{r}
# run me
clinical <- read.csv("data/clinical.csv") 
```

# earlier we used nested commands to execute multiple commands at once 

```{r}
# select patients with Stage IV cancer, then select only the race, ethnicity, and stage columns
race_stage <- select(filter(clinical, ajcc_pathologic_stage == 'Stage IV'), race, ethnicity, ajcc_pathologic_stage)

```
Take a look at the code above, what is the function filter doing?



What is the inner most nest (layer) of code?


#lets do the same task as above, but with pipes

# first, start off the pipe with the pipe operator, %>%, call this piped
# then write the first layer filtering for ajcc_pathologic_stage == "Stage IV",  use. to reference the clinical dataset
# then write the layer for selecting the race, ethnicity, and stage columns, use > to reference the clinical dataset you filtered above
```{r}

```
If you want to read more about piping,  https://magrittr.tidyverse.org/



# extract race, ethnicity, and age at diagnosis from cases born prior to 1930, filter before you select your columns, call this piped2
```{r}

```


# do the same as above, but this time select your columms before filtering rows, call this piped3
```{r}

```

## create a subsetted dataframe containing the column range tissue_or_organ_of_origin through total.number.Indels and contains RB1
## call this subsetted dataframe RB1_clinical

```{r}
RB1_clinical <- select(clinical,
                       tissue_or_organ_of_origin:total.number.Indels,
                       contains("RB1"))
```

# filter RB1_clinical for just the "Primary Tumor" samples types in "sample_type"
```{r}


```


# filter RB1_clinical for patients who have either "Del" or "Loss" in focal.deletion.in.RB1 OR "yes" in mutations.in.RB1 
# hint: we can use the "|" operator for OR. 
```{r}

```



# use the pipe operator to combine the selecting/filtering steps above
# you'll need a total of three expressions in your filtering step using the form: 

my_data_frame %>%
  fuction(inputs) %>%
  function(inputs)
  
```{r}

```


Great work!
