---
title: "8.1_piping_Exercises.Rmd"
output: html_document
date: "2024-01-24"
---

# Data: The Cancer Genome Atlas (TCGA) bladder cancer sequencing and patient characteristics

## Import packages

```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```

## Load the data

```{r}
clinical <- read.csv("data/clinical.csv")

# this is the selection procedure you used yesterday to select a subset of the columns
RB1_clinical <- select(clinical,
                       tissue_or_organ_of_origin:total.number.Indels,
                       contains("RB1"))
```


# Part I: Piping practice

#filter RB1_clinical for just the "Primary Tumor" samples types in "sample_type"
```{r}
# your code here...
```


#filter RB1_clinical for patients who have either "Del" or "Loss" in focal.deletion.in.RB1 OR "yes" in mutations.in.RB1 
hint: we can use the "|" operator for OR. 
```{r}

```


#use the pipe operator to combine the selecting/filtering steps above
```{r}

```




