---
title: "Data Manipulation"
author: "Sam Hart & Katherine Brower"
date: "2023-08-10"
output: html_document
---

__Under each heading below, type your code and run it. Feel free to add notes and comments to help you learn.__ 


# load library/package

```{r}
library(tidyverse)
```


# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# show categories in gender
```{r}
unique(clinical$gender) # remember, the $ extracts the column as a vector
```


# count number of individuals by gender
```{r}
gender_count <- 
```


# summarize average birth year by gender
```{r}
avg_YOB_by_gender <- 
```


# summarize average age at diagnosis (in days) by gender
```{r}
avg_age_days_by_gender <-
```


# summarize average, median, minimum, and maximum 
# age at diagnosis (in days) by gender. Don't forget to remove any NAs!

```{r}
clinical_summarized_by_gender <- 
```

# save clinical_summarized as a .csv file in our data directory

```{r}
# your code here...
```

# counting number of distinct diagnoses

```{r}
diagnosis_counts <- 
```

  
# get names of cancer diagnoses with more than 1 patient'

```{r}
specific_diagnoses <-
``` 

  
# extract data specifying primary diagnoses to keep

```{r}
clinical_reduced <- 
```

  
# confirm diagnoses present are the ones we want

```{r}
clinical_reduced$primary_diagnosis %>% 
  unique() # this is the same as unique(clinical$gender) but with a pipe
```


# same as above but without an intermediate object

```{r}
clinical_reduced <- clinical %>%
  filter(primary_diagnosis %in% #fill me in!)
```





