---
title: 'Lesson 6: Coding Exercises'
author: "Sam Hart & Katherine Brower"
date: "2023-08-10"
output: html_document
---

## Data: The Cancer Genome Atlas (TCGA) bladder cancer sequencing and patient
##       characteristics



## Part I: Counting and summarizing groups

## Instructions:
##  We are interested in how RB1--which is best known for contributing to eye
##  cancers--might contribute to bladder cancers. Let's continue with our expression
##  comparisons from last time!
##   - Group by "any.RB1.mut" and then determine the median, min, and max
##     of RB1_expr for each group using summarize.
##   - Group by cancer stage and then determine the median age of diagnosis *in years*
##     Hint: Remember to use View() or names() if you're not sure of the column
##      names to use here. Remember you can nest mathematical functions just like
##      R functions or the two together. :)


```{r}
# install.packages("tidyverse")
library(tidyverse) ## if this isn't working, be sure to run the above (uncomment first)
```

## Load the data

```{r}
clinical <- read_csv("data/clinical.csv")
```

## Continue with RB1-relevant subset of data

```{r}
RB1_clinical <- clinical %>%
  select(tissue_or_organ_of_origin:total.number.Indels, contains("RB1")) %>%
  filter(sample_type == "Primary Tumor") %>%
  mutate(across(c(focal.deletion.in.RB1,mutation.in.RB1), #find cols that contain "RB1"
                ~replace_na(., "NaN"))) %>%  #replace the NAs with strings
  mutate(any.RB1.mut = focal.deletion.in.RB1 == "Del" |
           focal.deletion.in.RB1 == "Loss" |
           mutation.in.RB1 == "yes"
  ) %>%
  mutate(RB1_expr = as.numeric(RB1_expr)) #make sure expression is numeric just in case
```

##   - Group by "any.RB1.mut" and then determine the median, min, and max
##     of RB1_expr for each group using summarize.

```{r}

```

##   - Group by cancer stage and then determine the median age of diagnosis *in years*
##     Hint: Remember to use View() or names() if you're not sure of the column
##      names to use here. Remember you can nest mathematical functions just like
##      R functions or the two together. :)

```{r}

```


## Part II: Saving data and more counting options

## Instructions:
##  - Save RB1_clinical to a new csv called "RB1.csv" in the directory data
##     Hint: If you're not sure what path to enter, use the read_csv statement
##      above as a template.
##  - Group by the column "any.RB1.mut" and count the number of patients with each cancer stage using tally()
##  - Filter race for Asian, Black, or White patients using the %in% operator;
##    then count the number of patients with each cancer stage using count()


##  - Save RB1_clinical to a new csv called "RB1.csv" in the directory data
##     Hint: If you're not sure what path to enter, use the read_csv statement
##      above as a template.
```{r}

```

##  - Group by the column "any.RB1.mut" and count the number of patients with each cancer stage using tally()
```{r}

```


##  - Filter race for Asian, Black, or White patients using the %in% operator;
##    then count the number of patients with each cancer stage using count()
```{r}

```



