---
title: "My first Rmd file"
output: html_document
date: "2024-01-12"
---

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: 
Today's date:
More notes:


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}
# insert code here

```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk.

We are going to use these Rmd files to write and save our code from now on!

# Rmd file practice/Intro to R Review
# 1) add 3 and 4 together, using the '+' operator, just like in math class
```{r}


```

Click on the run button to see the answer! It's the triangle button in the top right corner of the grey code chunk.

# 2a) assign the value 10 to the object weight_kg, by typing weight_kg <- 10
# 2b) what is weight_kg?
```{r}


```

Once you run the code chunk, you should see weight_kg show up in the Environment window!

# 3) convert kg to lb by multiplying weight_kg by 2.2
```{r}
```

# 4a) assign a new object called weight_lb with the value you got above
# 4b) what is the data type of weight_lb
```{r}


```

Once you run the code chunk, you should see weight_lb pop up in the Environment window.

# 5) re-assign weight_kg to 100
```{r}

```

You should see the value for weight_kg change in your Environment window. 

# 6) round 3.14159 to 3 digits
```{r}


```


# Vector Review!
# 7a) create a vector of your breakout group member's favorite animals called animal_party
# 7b) then add one animal to the front of the vector animal_party
# 7c) then add one marine animal to the end of the vector animal_party
# 7d) extract the 3rd animal and save it to an object called party_star
```{r}

```

```{r}

```

# Additional fun with the animal party!
# 8) What are 2 other things/functions you could do to add on or edit the vector animal party? Add them below!
```{r}

```


# Vector and Functions Review
# 9a) Create a vector called heights in your vector, add the heights of your breakout room group members in inches
# 9b) at the beginning of the vector, add your age
# 9c) at the end of the vector, add the height of one of someone from your family

```{r}

```


# 10) calculate the mean, median, maximum, and minimum of the vector heights
```{r}

```

# 11a) what kind of data is in the vector heights?
# 11b) Add one more thing/function that you can do with the vector heights! 
# remember you can use
```{r}

```


# Data Frames/tidyverse Review
# 12a) load the clinical dataset
```{r}

```

# 12b) get the structure of clinical
```{r}
str(clinical)
```

# 12c) load tidyverse
```{r}

```

# 12d) filter clinical  for rows that have an RB1 mutation (mutation.in.RB1 is 'yes'), assign this to filtered
```{r}

```

# subsetting dataframes rows and columns
# 13a) Filter clinical.csv to only have tumor samples
# 13b) Subset clinical.csv to 4 columns of your choice
```{r}



```

# creating vectors
# 14) Create a vector by setting it equal to a column in clinical.csv (choose a column made up of numbers, not words) 
# hint: to see the columns, run View(clinical.csv)
```{r}


```

# creating vectors
# 15) Create a vector by setting it equal to a column in clinical.csv (choose a column made up of numbers, not words) hint: to see the columns, run View(clinical.csv)
```{r}

```


# using na.rm and is.na
# 16a) Are there any NAs in your vector? 
# 16b) If so, remove them.
```{r}



```


Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  



