---
title: "My first Rmd file"
output: html_document
date: "2024-01-12"
---

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: 
Today's date:
More notes:


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}
# insert code here

```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk. For any notes within a code chunk, don't forget to use the # symbol!

We are going to use these Rmd files to write and save our code from now on!

# Rmd file practice/Intro to R Review
# 1) add 3 and 4 together, using the '+' operator, just like in math class
```{r}


```

Click on the run button to see the answer! It's the triangle button in the top right corner of the grey code chunk.

# 2) assign the value 15 to the object weight_kg
```{r}

```

Once you run the code chunk, you should see weight_kg show up in the Environment window!

# 3) convert kg to lb by multiplying weight_kg by 2.2
```{r}

```

# 4a) assign a new object called weight_lb with the value you got above
# 4b) what is the data type of weight_lb?
```{r}


```

Once you run the code chunk, you should see weight_lb pop up in the Environment window.

# 5) re-assign weight_kg to 100
```{r}

```

You should see the value for weight_kg change in your Environment window. 

# 6) round 3.14159 to 3 digits
```{r}


```


# Vector Review!
# 7a) create a vector of your breakout group member's favorite animals called animal_party
# 7b) then add one animal to the front of the vector animal_party
# 7c) then add one marine animal to the end of the vector animal_party
# 7d) extract the 3rd animal and save it to an object called party_star
# 7e) unfortunately the marine animal wasn't able to make it to the party, remove them from the vector
```{r}
#7a

```

```{r}
#7b

```

```{r}
# 7c 

```

```{r}
# 7d

```

```{r}
# 7e

```
# Additional fun with the animal party!
# 8) What are 2 other things/functions you could do to add on or edit the vector animal party? Add them below!
```{r}

```


# Vector and Functions Review
# 9a) Create a vector called heights in your vector, add the heights of your breakout room group members in inches
# 9b) at the beginning of the vector, add your age
# 9c) at the end of the vector, add the height of one of someone from your family

```{r}

```

# 10a) what kind of data is in the vector heights?
# 10b) Use 3 different functions with the vector heights (Ex. max, min, etc.)
# Remember you can always review the slides or previous worksheets to see what
# functions we've used before! Or use the ?function_name to try a new one
```{r}

```


# Data Frames/tidyverse Review
# Creating your own data frame
# 11a) create 3 different vectors with the length of 6:
# The first vector called boolean_vector containing TRUE or FALSE (boolean data)
# The second vector called favorite_foods containing your favorite foods
# The third vector called my_numbers containing 6, 10, 24, 61, 3, and 200

```{r}

```

# 11b) put boolean_vector, favorite_foods, and my_numbers into a dataframe
# called my_df. The column names should be boolean, character, and number.
```{r}


```

# 11c) look at the first few rows of my_df.
```{r}

```

# Utilizing the .csv file dataframe
# 12a) read in the pokemon.csv file as the object called pokemon_df
```{r}

```

# Write the answers to these questions below your code as a comment (don't forget to use #)
# 12b) how many rows and columns are there in pokemon_df?
# 12c) what is the name of the 12th column in the pokemon data frame?

```{r}

```

# 12d) what is the name of a column containing character data?
# 12e) what is the name of a column containing numeric or integers?
```{r}

```
# 13a) in your chosen column from 12e (column with numerics or integers):
# compute the following math with your chosen column containing numerics or integers: max, min, median, mean, standard deviation.
# Are there any NA values?
```{r}




```

# 13b) how many datapoints are less than the median?
# 13c) how many datapoints are more than or equal to the mean?
```{r}

```

# 13d) how many datapoints are within 30% of the median?
```{r}

```

# 14a) create a vector called pokemon_Total containing the column called Total in the pokemon_df
# 14b) how many pokemon have a total higher than 450?
```{r}

```

# 14c) create a vector called pokemon_Defense containing the column called Defense
# in the pokemon_df.
# challenge question 14d) which pokemon has the most Defense? (hint: use the which() function)
```{r}

```

# Debugging: Based on the given code, find the error! This could be using the wrong
# function, spelling, or using the wrong variable etc.! If you're stuck, try running the code 
# what is the error? What do you change to get the expected result?
```{r}
# 15a)  I want to find the max time of my vector called sprint_times. I am expecting 17.10 as the result.
sprint_times <- c(10.55, 17.10, 8.24, 11.45)
max(sprint_time)
```

```{r}
# 15b) I want to determine the median of my vector called years. I am expecting 2002.5 as the result
years <- c(2011, 2024, 2022, 1988, 1994)
mean(years)
```

```{r}
# 15c) I want to remove the NAs from my vector called route_numbers and calculate the mean.
# Hint: 3 errors! Expected result: 55
route_numbers <- s(25, 12, NA, NA, 99, 84)
mean(route_numbers, rm.na(route_numbers))
```

Excellent job!! 
Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  



