---
title: "1.0_intro_to_R"
output: html_document
date: "2024-01-12"
---

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: 
Today's date:
More notes:


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}
#insert code here

```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk.

We are going to use these Rmd files to write and save our code from now on!

Let's practice using an Rmd file

#add 3 and 4 together, using the '+' operator, just like in math class
```{r}


```

Click on the run button to see the answer!


#assign the value 10 to the object weight_kg, by typing weight_kg <- 10
```{r}

```

Once you run the code chunk, you should see weight_kg show up in the Environment window!


#convert kg to lb by multiplying weight_kg by 2.2
```{r}

```

#assign a new object called weight_lb with the value you got above
```{r}

```

Once you run the code chunk, you should see weight_lb pop up in the Environment window.

#re-assign weight_kg to 100
```{r}

```

You should see teh value for weight_kg change in your Environment window. 

#round 3.14159 to 3 digits
```{r}

```

#make a vector called practice_vector with the values (1, 3, 5, 7, 10)
```{r}

```

#find the median of practice_vector
```{r}

```

#load the clinical dataset
```{r}

```

#get the structure of clinical
```{r}

```

#load tidyverse
```{r}

```

#filter clinical  for rows that have an RB1 mutation (mutation.in.RB1 is 'yes'), assign this to filtered
```{r}

```


Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  



