# 1.0 Intro to R

Name: 

Date:

---

The gray box below is called a **code chunk**. We use it to write code.

```{r}

```

To run your code chunk, you can the green play button (Run current chunk) on the top right corner of the chunk. Let's try it here:

```{r}
# We use the "#" symbol to specify a line that R will not run as code. You can use it to write notes! Example below:

# This line of code prints the phrase Hello word
print("Hello World")
#print("This should be not printed")
```



The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

---

## Using R as a calculator

```{r}
# Add two values
12 + 4 
```

```{r}
5 + 4 * 6 - 2
```

```{r}
(5 + 4) * 6 - 2
```

```{r}
# Add your own equation for R to compute


```

---

## Assigning variables

We use the `<-` operator to store a value to a variable.

```{r}
x <- 15
```

To see what `x` is, we call it by typing it out:

```{r}
x
```

Alternatively, we can look in the `Environment` pane to see all objects and its values.

You can also use `=` to assign a variable. `=` works the same as `<-`.

```{r}
y = 16
```


Here is an example for assigning objects and using the objects to form a new object:

```{r}
apples <- 5
oranges <- 6

total_fruits <- apples + oranges
total_fruits
```


Let's create more fruit variables and assign them with numbers. Then, recalculate the number of total fruits. 
```{r}

```

-----------

## Data types

The most common data types we will work with are:

-   **Character**, also called **strings**: Contains a series of letters or symbols enclosed in `"` or `'`. Examples: `"Hello!"`, `"coding4%"`, `"good morning 52"`
-   **Numeric**: Contains any type of number data. Example: `4.1503`
-   **Integer**: Contains number data, but do not have decimal points. Example: `5`
-   **Logical**: Contains only `TRUE` or `FALSE`. It can be also represented as `T` or `F`, respectively.

From last time, we use the `=` operator to assign values to variables.

```{r}
# Assign an integer or numeric to the variable myAge
myAge = 
```

```{r}
# Declare variable `myName` to store a character type of your name. Don't forget the quotation marks!
myName = 
```

```{r}
# Assign a logical if you have, or have not, visited the Space Needle.  
visitedSpaceNeedle = 
```

If you assigned `myName`, `myAge` and `visitedSpaceNeedle` correctly, then if you run this code block, it should say a correct statement.

```{r}
# You do not need to modify this code chunk. 
print(paste0("My name is ", myName))
print(paste0("I am ", myAge, " years old"))
if(visitedSpaceNeedle == T) {print("I have visited the Space Needle.")}
if(visitedSpaceNeedle == F) {print("I have not visited the Space Needle.")}
```

