---
title: "1.1 R Practice"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
```

# 1.1 R Practice

We will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

------------------------------------------------------------------------

## Recap

We can make variables and use `=` or `<-` to assign a value to the variable.

```{r}
apples = 10
potatoes <- 3
```

There are different data types in R. The most common ones are 1) strings, 2) integers, 3) numeric, 4) logical.

This is an example of a string:

```{r}
city = "Seattle"
```

It does not make sense if you do math operations on strings:

```{r}
city + 4
```

------------------------------------------------------------------------

## Example 1: Using R as a calculator

Let start with the variable `myAgeYears`. Assign the variable `myAgeYears` with your age in years.

```{r}
myAgeYears = 
```

Convert `myAgeYears` to days. Assign this value to `myAgeDays`.

```{r}
myAgeDays = 
```

Calculate your age in seconds. Assign this value to `myAgeSeconds`.

```{r}

```

------------------------------------------------------------------------

## Example 2: Heart beat calculations

R can also be used for more complicated calculations, let's calculate how many times a heart beats in a year.

A normal resting heart rate for adults ranges from 60 - 100 beats per minute (bpm).

First, let's make an object called `lower_range_bpm` and assign it 60.
```{r}

```

Now let's do the same for the upper range of heartbeats. Assign the object `upper_range_bpm` with the upper range of heart beats.
```{r}

```


Now we need to calculate how many minutes in a year, let's start by multiplying 60 minutes by 24 hours to get the number of minutes in a day and assign it to `mins_in_day`. 
```{r}
mins_in_day 
```

