# 10.0 Intro to plots: Bar plots and Histograms

We will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Load data and libraries

```{r}
library(tidyverse)
library(palmerpenguins)
penguins <- penguins
```

## 1. My first plot

We will use the `ggplot` package. The major components to a `ggplot` graph are:

-   Data (the information we want to show)

-   Aesthetics (the axes)

-   Geometry (the shapes, the plot type)

-   Themes (what everything looks like)

    ```{r}
    ggplot(penguins) + aes(x = bill_length_mm) + geom_histogram() 
    ```

We can add a preset theme to our plot in a similar grammer structure.

```{r}
ggplot(penguins) + aes(x = bill_length_mm) + geom_histogram() + theme_bw()
```

Here is a list of themes readily available:

-   `theme_gray()` The signature ggplot2 theme with a grey background and white gridlines, designed to put the data forward yet make comparisons easy.

-   `theme_bw()` The classic dark-on-light ggplot2 theme. May work better for presentations displayed with a projector.

-   `theme_linedraw()` A theme with only black lines of various widths on white backgrounds, reminiscent of a line drawing. Serves a purpose similar to `theme_bw()`.

-   `theme_light()` A theme similar to `theme_linedraw()` but with light grey lines and axes, to direct more attention towards the data.

-   `theme_dark()` The dark cousin of `theme_light()`, with similar line sizes but a dark background. Useful to make thin coloured lines pop out.

-   `theme_minimal()` A minimalistic theme with no background annotations.

-   `theme_classic()` A classic-looking theme, with x and y axis lines and no gridlines.

-   `theme_void()` A completely empty theme.

Try a different theme here:

```{r}
ggplot(penguins) + aes(x = bill_length_mm) + geom_histogram() 
```

## 2. Bar plots

Barplots plots categorical data and makes numerical comparisons, usually counts, between categories.

We will use the clinical dataset for more plotting examples.

```{r}
clinical <- read_csv("clinical.csv", show_col_types = FALSE)
```

Plot categorical data in a bar plot for the `mutation.in.TP53` data. For a bar plot, we use `geom_bar()` as the geometry type.

```{r}

```

## 3. Histogram

Histogram plots the range of a set of numerical values.

Plot a histrogram showing the range of `height`. For a histogram, we use `geom_histogram()` as the geometry type.

```{r}


```

Changing the number of bins: `bins=10`

```{r}

```

Changing the width of the bins: `binwidth=10`

```{r}

```
