# 11.0 Intro to plots: Scatter and Box plots

We will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Load data and libraries

```{r}
library(tidyverse)
library(palmerpenguins)
penguins <- penguins
clinical <- read_csv("clinical.csv", show_col_types = FALSE)
```

## Scatter plots

Let's make a scatter plot to compare the number of cigarettes per day (`cigarettes_per_day` column) with the number of ERCC2 mutations (`count.ERCC2.signature.mutations` column.)

```{r}

```

We can alter the coordinate range to show a specific region using `coord_cartesian(xlim = c(0, 10), ylim = c(0, 400))`

```{r}

```

We can use pipes to remove NAs to clean the plot. Recall this from the piping lesson:

```{r}
clinical %>%
  select(barcode, cigarettes_per_day, count.ERCC2.signature.mutations) %>%
  drop_na
```

This input can go directly into `data=`

```{r}

```

## Box plots

Let's make a scatter plot to compare `Current smoker`and `Lifetime Non-Smoker` (both in `Tobacco.smoking.history` column) with the number of ERCC2 mutations (`count.ERCC2.signature.mutations` column.)

First let's filter the data to just our data of interest:

```{r}
barcode_TSH_ERCC2_count_data <- clinical %>% 

```

Plot the filtered dataset:

```{r}

```
