# 2.0 Basic Data Types

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Data types

The most common data types we will work with are:
  
  - **Character**, also called **strings**: Contains a series of letters or symbols enclosed in `"` or `'`. Examples: `"Hello!"`, `coding4%`, `good morning 52`
  - **Numeric**: Contains any type of number data. Example: `4.1503`
  - **Integer**: Contains number data, but do not have decimal points. Example: `5`
  - **Logical**: Contains only `TRUE` or `FALSE`. It can be also represented as `T` or `F`, respectively. 


From last time, we use the `<-` operator to assign values to variables.

```{r}
# Assign an integer or numeric to the variable myAge
myAge <- 
```

```{r}
# Declare variable `myName` to store a character type of your name. Don't forget the quotation marks!
myName <-
```

```{r}
# Assign a logical if you are, or are not, applying to college this upcoming cycle.  
applyingToCollege <-
```

If you assigned `myName` correctly, then if you run this code block, it should say a correct statement.
```{r}
# You do not need to modify this code chunk. 
print(paste0("My name is ", myName))
print(paste0("I am ", myAge, " years old"))
if(applyingToCollege == T) {print("I am applying to college this summer.")}
if(applyingToCollege == F) {print("I am not applying to college this summer.")}
```