# 3.0 Introduction to Vectors

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```


## Defining a vector

Recall from last lesson, we assign single values inside a variable like this:
```{r}
my_age <- 16
```

If we want store multiple objects, we assign them to an object called a **vector**. 

To do this in R, we put the values in this `c()` function. The `c()` function is used to **c**ombine values.
```{r}
ages <- c(50,55,60,65)
ages
```

We can use the `c()` function to add more values is an existing vector.
```{r}
more_ages <- c(ages, 70)
more_ages
```


## Indexing
Each element in a vector has an associated **index**, or **position**, that we can use as a reference. The **index** tells us the **position** of the element in the vector. 

Take the following vector as an example:
```{r}
organs <- c("lungs", "heart", "kidney", "brain")
```

If we want to access the 3rd value, we use square brackets `[]` with the vector like this:
```{r}
organs[3]
```


We can also use indexing to extract multiple values from the vector. This is called **subsetting** a vector.

To get the first **through** thrid elements use the `:` symbol like this:
```{r}
organs[1:3]
```

To get the **only first and third**, use a vector with 1 and 3 or `c(1,3)`
```{r}
organs[c(1,3)]
```

Alternatively, it can be written more defined like this:
```{r}
index <- c(1,3)

organs[index]
```