---
title: "5.0_Manipulating_Vectors_worksheet"
author: "Katherine Brower (kbrower)"
date: "2024-06-11"
output: html_document
---
# "5.0 Manipulating Vectors

We will start with a fresh, clean environment. 
The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Manipulating Vectors
Vectors are very useful for storing multiple values. 

We can create a vector using 
vector_name <- c(value 1, value 2, value 3 ...)

```{r}
# Create a vector called favorite_numbers and store 6 values

favorite_numbers <- c(13, 34, 86, 104, 22, 96)

```

We can subset values within a vector using vector_name[condition]

Examples: vector[index] or vector[index1, index2] or vector[index1:index2]

```{r}
# Extract values that are higher than 50 in favorite_numbers 

favorite_numbers[favorite_numbers > 50]

```

```{r}
# Extract values that are not higher than 40 in favorite_numbers

favorite_numbers[!(favorite_numbers>40)]

```

We can also use multiple conditions to extract values!

- **&** inclusively combine conditions (AND)

- **|** exclusively combine contidions (OR)

```{r}
# Extract values that are greater than 4 AND less than 55

favorite_numbers[favorite_numbers > 4 & favorite_numbers < 55]
```


```{r}
# Extract values that are greater than 40 OR less than 90
favorite_numbers[(favorite_numbers > 40) | (favorite_numbers < 90)]
```

Sometimes our datasets have missing data! This value will be written as NA.

We need to remove those NA values because R cannot compute mean or max with NA values.

The argument to do this is: na.rm = TRUE

```{r}
# create a vector of 6 values and 2 missing data points
numbers <- c(4, 10, 7, NA, 35, 86, 32, NA)

# calculate the mean and remove the NA
mean(numbers, na.rm = TRUE)

# calculate the max and remove the NA
max(numbers, na.rm = TRUE)
```

If you are uncertain if there are missing data points, we can use the is.na() function

```{r}
# create a vector of 4 values and  missing data
scores <- c(10, 58, NA, 35, 29, NA)

# use the is.na function to determine which data points are missing
is.na(scores)

```

```{r}
#we can subset the vector based on that condition
scores[is.na(scores)]
```