# 13.0 Plots 2.0: Effective data visualization

We will start with a fresh, clean environment. The code chunk below clears your environment if you want to reset your work in interactive mode. Or use the broom button on the top right corner.

```{r}
rm(list = ls())
```

## Load data and libraries

```{r}
library(tidyverse) 
library(palmerpenguins) 
penguins <- penguins
clinical <- read_csv("clinical.csv", show_col_types = FALSE)
```

We'll work primarily this data:

```{r}
barcode_TSH_ERCC2_count_data <- clinical %>% 
  select(barcode, Tobacco.smoking.history, count.ERCC2.signature.mutations) %>%
  filter(!is.na(count.ERCC2.signature.mutations)) %>% # remove NAs
  filter(Tobacco.smoking.history %in% c("Current smoker", "Lifelong Non-smoker")) # filter
```

Starting plot

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color=Tobacco.smoking.history))
```

## General changes and additions

### Titles and Axis Labels

We can add text labels to the existing ggplot formula:

-   For titles use `ggtitle("title")`

-   For axis labels use `labs(y = "y_axis_name", x = "x_axis_name")`

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color=Tobacco.smoking.history)) + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history")
```

### Removing legends

To remove legends, we can specify **`theme(legend.position = “none”)`**

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color=Tobacco.smoking.history)) + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none")

```

### Color with custom hex colors

We can change the **color** the jitter plot by a category with the aesthetic parameter `geom_jitter(aes(color=Tobacco.smoking.history))`.

Then we can set custom colors using `scale_color_manual` with either:

-   A color from the default R color list(<https://www.datanovia.com/en/blog/awesome-list-of-657-r-color-names/>). Example: `"red"`, `"gray80"`, `"deepskyblue"`

-   hexidecimal code (hex code, <https://htmlcolorcodes.com/>). Example: `"#fdae61"` , "**`#C625A1"`**, `"#535549"`

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color=Tobacco.smoking.history)) + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none") + 
  scale_color_manual(values = c("#fdae61", "#15BADC"))


```

### Filling shapes with color

Both **`color`** and **`fill`** have attributes that can be colored.

Generally, **`fill`** defines for shapes to be filled. **`color`** defines outlines (stroke, lines) of shapes. Technically, points are strokes, not shapes.

For filling shapes, we will use `fill = Tobacco.smoking.history` in the aesthetics and set the fill colors with `scale_fill_manual(values = c("#fdae61", "#C625A1"))`

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations, fill = Tobacco.smoking.history) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter() + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none") + 
  scale_color_manual(values = c("#fdae61", "#15BADC"))
```

### Transparency

We can change the transparency of `fill` or `color` by adding `alpha=0.5` in the aesthetics. Alpha can be values from 0 to 1.

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations, fill = Tobacco.smoking.history, alpha=0.2) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter() + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none") + 
  scale_color_manual(values = c("#fdae61", "#15BADC"))
```

### Modifying text color and size

We can change the text color and size of the axis and labels with

```         
  theme(axis.text.y = element_text(color = "blue", size = 10),
        axis.text.x = element_text(color = "blue", size = 10, angle = 45, hjust = 1),
        axis.title.y = element_text(color = "blue", size = 12),
        axis.title.x = element_text(color = "blue", size = 12))
```

Similarly, `color` can accept hex code colors, and the `angle` can be set. `hjust=1` stands for horizontal adjustment to make text align at the top (`=1`). `hjust=0` aligns the text at the bottom.

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations, fill = Tobacco.smoking.history, alpha=0.2) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter() + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none") + 
  scale_color_manual(values = c("#fdae61", "#15BADC")) + 
   theme(axis.text.y = element_text(color = "blue", size = 10),
        axis.text.x = element_text(color = "blue", size = 10, angle = 45, hjust = 1),
        axis.title.y = element_text(color = "blue", size = 12),
        axis.title.x = element_text(color = "blue", size = 12))
```

### Adding gridlines

We can add gridlines by altering the `theme` with **`theme(panel.grid.major = element_line(colour = "black", linetype="dashed",linewidth=0.25))`**

See the `ggplot theme` documentation (<https://ggplot2.tidyverse.org/reference/theme.html>) for more ways to modify plots.

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, y = count.ERCC2.signature.mutations, fill = Tobacco.smoking.history, alpha=0.2) + 
  geom_boxplot(outlier.shape = NA) + coord_cartesian(ylim = c(0, 125)) +
  geom_jitter() + 
  ggtitle("Tobacco smoking history vs ERCC2 mutation counts") + 
  labs(y = "ERCC2 mutation counts", x = "Tobacco smoking history") + 
  theme(legend.position = "none") + 
  scale_color_manual(values = c("#fdae61", "#15BADC")) + 
   theme(axis.text.y = element_text(color = "blue", size = 10),
        axis.text.x = element_text(color = "blue", size = 10, angle = 45, hjust = 1),
        axis.title.y = element_text(color = "blue", size = 12),
        axis.title.x = element_text(color = "blue", size = 12)) + 
  theme(panel.grid.major = element_line(color = "black", linetype="dashed", linewidth=0.25))
```

## Plot specific changes (and some more general additions)

### Bar plot

Let's start with this bar plot. Notice that you can set `y=` to make a horizontal bar plot.

```{r}
ggplot(data = clinical) + 
  aes(y = ajcc_pathologic_stage) + 
  geom_bar()
```

We can change the bar width, color, and fill by adding it in the geometry `geom_bar()` like so:

```         
geom_bar(width = 0.2, color = "red", fill = "orange")
```

```{r}
ggplot(data = clinical) + 
  aes(y = ajcc_pathologic_stage) + 
  geom_bar(width = 0.6, color = "red", fill = "orange")
```

We can add additional variables to go in the plots by assigning `fill` to another variable. Let's try coloring the bars based on **`tissue_or_organ_of_origin`**.

```{r}
ggplot(data = clinical) + 
  aes(y = ajcc_pathologic_stage, fill= tissue_or_organ_of_origin) + 
  geom_bar()
```

Alternatively, stacked bars can be hard to read, so bars side-by-side is another approach to visualize bar charts. To have side-by-side bars, add `position="dodge"` in the `geom_bar()` geometry.

```{r}
ggplot(data = clinical) + 
  aes(y = ajcc_pathologic_stage, fill= tissue_or_organ_of_origin) + 
  geom_bar(position="dodge")
```

### Histogram

Starting plot

```{r}
ggplot(data = clinical) + 
  aes(x = height) + 
  geom_histogram(bins=20)
```

Similar as bar charts, `fill` and `color` can be used to color the individual bars in the historgram

```         
geom_histogram(bins=20, fill = "white", color = "black")
```

```{r}
ggplot(data = clinical) + 
  aes(x = height) + 
  geom_histogram(bins=20, fill = "white", color="black")
```

We also can use the `fill` to display an additional element of data. In this case, let's try `fill=gender` in the aesthetics `aes(x = height, fill = gender)`.

```{r}
ggplot(data = clinical) + 
  aes(x = height, fill = gender) + 
  geom_histogram(bins=20)
```

To avoid stacked histograms, we add `position="identity"` in `geom_histogram()`. This makes the histograms overlap. Don't forget to alter the transparency to see the overlaps!

```         
geom_histogram(bins=20, position="identity")
```

```{r}
ggplot(data = clinical) + 
  aes(x = height, fill = gender, alpha=0.5) + 
  geom_histogram(bins=20, position="identity")
```

"

### Aside: `facet_wrap` can be used for all plot types

An alternative method to display different groups is to use `facet_wrap()`. If we can to group histograms by `gender`, we can add it the `ggplot` formula like this: `+ facet_wrap(~gender)`. Note: `facet_wrap()` can be used for any plot type, not just histograms.

```{r}
ggplot(data = clinical) + 
  aes(x = height, fill = gender) + 
  facet_wrap(~gender) + 
  geom_histogram(bins=20)
```

You can group any number of categories using `facet_wrap()`. Seperate the groups using the `~` symbol, like this: `facet_wrap(group1~group2)`. Try grouping `gender` with one of the following:

-   `race`

-   `Vital.status`

-   `mutation.in.TP53`

```{r}

```

You can change add additional details to `facet_wrap` like number of columns or row, and labels. Change the number of columns that would make the most sense for comparing groups.

```         
facet_wrap(mutation.in.TP53~gender, ncol=2, labeller = "label_both")
```

```{r}

ggplot(data = clinical) + 
  aes(x = height, fill = gender) + 
  facet_wrap(mutation.in.TP53~gender, ncol=2, labeller = "label_both") +
  geom_histogram(bins=20)
```

Lastly, we can add vertical (or horizontal) lines to indicate key statistics like mean or median. One way to accomplish this is group the data with `group_by`, then introduce a new column with `mutate` of the statistic.

```         
  geom_vline(data = clinical %>% 
                      group_by(mutation.in.TP53, gender) %>% 
                      mutate(mean_height = mean(height, na.rm = TRUE)), 
             aes(xintercept = mean_height), color="blue", linetype = "dashed")
```

```{r}
ggplot(data = clinical) + 
  aes(x = height, fill = gender) + 
  facet_wrap(mutation.in.TP53~gender, ncol=2, labeller = "label_both") +
  geom_histogram(bins=20) + 
  geom_vline(data = clinical %>% 
                      group_by(mutation.in.TP53, gender) %>% 
                      mutate(mean_height = mean(height, na.rm = TRUE)), 
             aes(xintercept = mean_height), color="blue", linetype = "dashed")
```

### Scatterplots

Starting plot

```{r}
ggplot(data = clinical) + 
  aes(x = height, y = bmi) +
  geom_point()
  
```

Similar to other plots, we can change the color of the points using `geom_point(color="pink")` in the geometry.

We can add a trend line using `+ geom_smooth()`

```{r}
ggplot(data = clinical) + 
  aes(x = height, y = bmi) +
  geom_point() + 
  geom_smooth()
```

We can use shapes as points as an additional dimension and a great alternative for colorblind accessibility.

```{r}
ggplot(data = clinical) + 
  aes(x = height, y = bmi, shape=gender, color=gender) +
  geom_point()
```

### Box plots and violin plots

Starting plot

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, 
      y = count.ERCC2.signature.mutations,
      fill = Tobacco.smoking.history) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = c(0, 125)) +
  ggtitle("Number of ERCC2 mutations vs. Tobacco smoking history") + 
  labs(y = "Number of ERCC2 mutations", 
       x = "Tobacco smoking history") + 
  scale_fill_manual(values = c("#fdae61", "#C625A1"))

```

```{r}
ggplot(data = barcode_TSH_ERCC2_count_data) + 
  aes(x = Tobacco.smoking.history, 
      y = count.ERCC2.signature.mutations,
      fill = Tobacco.smoking.history) + 
  geom_violin() + 
  coord_cartesian(ylim = c(0, 125)) +
  ggtitle("Number of ERCC2 mutations vs. Tobacco smoking history") + 
  labs(y = "Number of ERCC2 mutations", 
       x = "Tobacco smoking history") + 
  scale_fill_manual(values = c("#fdae61", "#C625A1"))

ggsave("violin_plot.png")
```

## Exit Ticket

Turn in your plot from today as your Exit Ticket.

Here's the link to today's Exit Ticket: <https://forms.gle/S3JGSPdSghdHhjcA9>
