---
title: "Plots 2.0 Exercises KEY"
author: "sfhart"
date: "7/17/2022"
output: html_document
---

# load library/package
```{r}
library(tidyverse)
```

# import data
```{r}
clinical <- read_csv("data/clinical.csv")
```

# boxplot from lesson 7a
```{r}
d_plot <- clinical %>% 
  select(barcode, Tobacco.smoking.history, count.ERCC2.signature.mutations) %>% 
  filter(!is.na(count.ERCC2.signature.mutations)) %>% 
  filter(Tobacco.smoking.history %in% c("Current smoker", "Lifelong Non-smoker"))

L7A_boxplot <- ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "blue")

L7A_boxplot
```


# set boxplot color = "red"
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           color = "red")) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) 
  geom_jitter(color = "blue")
```

# set boxplot fill = "red"
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = "red")) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "blue")
```

# set boxplot fill to be colored by Tobacco smoking history
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "blue")
```

# specify boxplot fill colors
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(color = "blue") +
  scale_fill_manual(limits = c("Current smoker", "Lifelong Non-smoker"), # category names
                    values = c("orange", "blue")) # corresponding colors
```

# specify boxplot fill and jitter points colors
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color = Tobacco.smoking.history)) + # specify data column
  scale_fill_manual(limits = c("Current smoker", "Lifelong Non-smoker"),
                    values = c("orange", "blue"))
```

# use hex color codes
# choose your own or use "#fdae61" and "#abd9e9" 

```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color = Tobacco.smoking.history)) +
  scale_fill_manual(limits = c("Current smoker", "Lifelong Non-smoker"),
                    values = c("#fdae61", "#abd9e9")) # hex color codes
```

# add transparency to jittered points
```{r}
ggplot(data = d_plot,
       aes(x = Tobacco.smoking.history,
           y = count.ERCC2.signature.mutations,
           fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color = Tobacco.smoking.history), alpha = 0.5) + # add transparency
  scale_fill_manual(limits = c("Current smoker", "Lifelong Non-smoker"),
                    values = c("#fdae61", "#abd9e9"))
```

# save plot to an object p
# don't forget to add a title and axis labels!
```{r}
# Alpha =0.5
p <- ggplot(data = d_plot,
            aes(x = Tobacco.smoking.history,
                y = count.ERCC2.signature.mutations,
                fill = Tobacco.smoking.history)) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0, 125)) +
  geom_jitter(aes(color = Tobacco.smoking.history), alpha = 0.5) +
  scale_fill_manual(limits = c("Current smoker", "Lifelong Non-smoker"),
                    values = c("#fdae61", "#abd9e9")) +
 labs(title = "Tobacco smoking history vs. Number of Signature ERCC2 Mutations",
       x = "Tobacco smoking history",
       y = "Number of ERCC2 signature mutations")

p
```

# modify plot area and gridlines, use a new theme!
```{r}
p +
  theme_bw() # using a pre-set theme
```
  
# modify plot text features with theme() to make it more readable!
# try color = black and size = 7
# you can put the axis at an angle if it makes it easier to read!
```{r}
p +
  theme_bw() + 
  theme(axis.text.y = element_text(color = "black", size = 7),
        axis.text.x = element_text(size = 7, color = "black", angle = 45, hjust = 1),
        axis.title.y = element_text(color = "black", size = 7),
        axis.title.x = element_text(color = "black", size = 7))
```

# remove legend
```{r}
p +
  theme_bw() + 
  theme(axis.text.y = element_text(color = "black", size = 7),
        axis.text.x = element_text(size = 7, color = "black", angle = 45, hjust = 1),
        axis.title.y = element_text(color = "black", size = 7),
        axis.title.x = element_text(color = "black", size = 7),
        legend.position = "none")
```

# fix axis titles
```{r}
p +
  theme_bw() + 
  theme(axis.text.y = element_text(color = "black", size = 7),
        axis.text.x = element_text(size = 7, color = "black", angle = 45, hjust = 1),
        axis.title.y = element_text(color = "black", size = 7),
        axis.title.x = element_text(color = "black", size = 7),
        legend.position = "none") +
  labs(x = "Tobacco smoking history",
       y = "Number of ERCC2 signature mutations")
```

# create final_plot object
```{r}
final_plot <- p +
  theme_bw() + 
  theme(axis.text.y = element_text(color = "black", size = 7),
        axis.text.x = element_text(size = 7, color = "black", angle = 45, hjust = 1),
        axis.title.y = element_text(color = "black", size = 7),
        axis.title.x = element_text(color = "black", size = 7),
        legend.position = "none") +
  labs(title = "Tobacco smoking history vs. Number of Signature ERCC2 Mutations",
       x = "Tobacco smoking history",
       y = "Number of ERCC2 signature mutations")

final_plot
```
 
# save plot to plot directory
# name it something descriptive!
```{r}
ggsave(filename = "plots/boxplot.pdf",
       plot = final_plot,
       width = 1.5, 
       height = 3) 
```



