---
title: "7.1_tidyverse_Exercises_KEY.Rmd"
output: html_document
date: "2024-01-23"
---

welcome to the tidyverse Exercises!

we are going to continue working with the clinical data frame using tidyverse!

#before we begin, you should have tidyverse installed (skip this if you)
```{r}
install.packages("tidyverse")
```

#load tidyverse
```{r}
library(tidyverse)
```

#load the clinical data set we have been working with
```{r}
clinical <- read.csv('data/clinical.csv')
```

#find the structure of the data set
```{r}
str(clinical)
```

today we are interested in looking at the RB1 gene
the RB1 gene is known to contribute to eye cancers, but we want to see if it might contribute to bladder cancers

#look at all columns that contain RB1, extract it and assign it to RB1_all
```{r}
RB1_all <- select(clinical, contains('RB1'))
```

#view this data frame
```{r}
View(RB1_all)
```

Which one of these columns that contain RB1 contains data as yes/no?


#extract the columns tissue_or_organ_of_origin THROUGH total.number.Indels, assign this to cancer_clinical
```{r}
cancer_clinical <- select(clinical, tissue_or_organ_of_origin:total.number.Indels)
```

now add those parameters together to make 1 data frame

#select the columns tissue_or_organ_of_origin THROUGH total.number.Indels AND all columns that contain RB1, assign this to RB1_clinical
```{r}
RB1_clinical <- select(clinical, tissue_or_organ_of_origin:total.number.Indels, contains('RB1'))
```

#take a look!
```{r}
View(RB1_clinical)
```

Which column contains information on the stage of cancers?
ajcc_pathologic_stage

now that we have the column we want, lets filter the rows to get the data we want to analyze

how many rows are in RB1_clinical?
433

#filter RB1_clinical to show just the rows taht have 'Primary Tumor' in sample_type, assign this to filter1
```{r}
filter1 <- filter(RB1_clinical, sample_type == "Primary Tumor")
```

#take a look at the structure of filter1
```{r}
str(filter1)
```

How many rows are in filter1?
414


#filter RB1_clinical for patients who have either 'Del' or 'Loss' in the focal.deletion.RB1, assign this to filter2
```{r}
filter2 <- filter(RB1_clinical, focal.deletion.in.RB1 == 'Del' | focal.deletion.in.RB1 == "Loss")
```

#print filter2
```{r}
filter2
```

How many rows are in filter2?
64


#now lets filter for rows that have an RB1 mutation (mutation.in.RB1 is 'yes'), assign this to filter3
```{r}
filter3 <- filter(RB1_clinical, mutation.in.RB1 == 'yes')
```

#take a look at filter3
```{r}
View(filter3)
```

How many rows are in filter3?
70

remember RB1_clinical is already narrowed down to show related columns
and what we are interested in looking at is the data of patients who have deleted or lost RB1 or has a mutation in RB1


#lets combine the filters from above to filter for RB1_clinical patients who have either 'Del' OR 'Loss' in focal.deletion.in.RB1 OR 'yes' in mutation.in.RB1, assign this to filter_all
```{r}
filter_all <-  filter(RB1_clinical, focal.deletion.in.RB1 == "Del" |
  focal.deletion.in.RB1 == "Loss" | mutation.in.RB1 == "yes")
```

#open filter_all
```{r}
View(filter_all)
```

How many rows are in filter_all?
114 

great work subsetting your dataset, now we have the data we are interested in!

