---
title: "My first Rmd file"
output: html_document
date: "2024-01-12"
---

This is a R Markdown document.

You can take notes just like you can in word or google docs. 

Try writing a few notes!

Your name: Katherine Brower
Today's date: 8/7/24
More notes: example key


R Markdown documents can also contain code chunks. The grey area below is a code chunk:

```{r}
# I am a code chunk, when there is a # at the beginning of the code, this will
# be read as a comment.

```

Code written in a code chunk can be run just like you wrote it in the Console pane below.
To run the code in the code chunk, click on the play button in the top right corner
of each code chunk. For any notes within a code chunk, don't forget to use the # symbol!

We are going to use these Rmd files to write and save our code from now on!

# Rmd file practice/Intro to R Review
# 1) add 3 and 4 together, using the '+' operator, just like in math class
```{r}
3 + 4 

```

Click on the run button to see the answer! It's the triangle button in the top right corner of the grey code chunk.

# 2) assign the value 15 to the object weight_kg
```{r}
weight_kg <- 15
```

Once you run the code chunk, you should see weight_kg show up in the Environment window!

# 3) convert kg to lb by multiplying weight_kg by 2.2
```{r}
weight_kg * 2.2
```

# 4a) assign a new object called weight_lb with the value you got above
# 4b) what is the data type of weight_lb?
```{r}
weight_lb <- weight_kg * 2.2

```

Once you run the code chunk, you should see weight_lb pop up in the Environment window.

# 5) re-assign weight_kg to 100
```{r}

weight_kg <- 100
```

You should see the value for weight_kg change in your Environment window. 

# 6) round 3.14159 to 3 digits
```{r}
round(3.14159, 3)

```


# Vector Review!
# 7a) create a vector of your breakout group member's favorite animals called animal_party
# 7b) then add one animal to the front of the vector animal_party
# 7c) then add one marine animal to the end of the vector animal_party
# 7d) extract the 3rd animal and save it to an object called party_star
```{r}
#7a
animal_party <- c("dog", "penguin", "kakapo", "cat", "bear")

```

```{r}
#7b
animal_party <- c("octopus", animal_party)
```

```{r}
#7c
animal_party <- c(animal_party, "starfish")
animal_party[7]
```

```{r}
#7d 
party_star <- animal_party[3]
party_star
```

```{r}
# 7e 
animal_party <- animal_party[-length(animal_party)]
animal_party
```

```{r}
#example if we wanted to remove the 5th animal
animal_party <- animal_party[-5]
animal_party
```
# Addi tional fun with the animal party!
# 8) What are 2 other things/functions you could do to add on or edit the vector animal party? Add them below!
```{r}
# example 1 adding an animal in the 4th index/spot (challenge, there are)
# other functions from packages that we will use in future lessons that
# will do this same action!
animal_party <- c(animal_party[1:3], "rabbit", animal_party[4:length(animal_party)])

# example 2: we had some late guests, let's add them to the party!
# adding a new vector called late_arrival to the end of the animal_party vector

late_arrival <- c("sloth", "elephant", "giraffe", "tortoise")
```

```{r}
# combined animal_party with the late_arrival
animal_party <- c(animal_party, late_arrival)
animal_party
```


# Vector and Functions Review
# 9a) Create a vector called heights in your vector, add the heights of your breakout room group members in inches
# 9b) at the beginning of the vector, add your age
# 9c) at the end of the vector, add the height of one of someone from your family

```{r}
#9a
heights <- c(65, 58, 70, 68)

#9b
heights <- c(28, heights)

#9c
heights <- c(heights, 72)
```

# 10a) what kind of data is in the vector heights?
# 10b) Use 3 different functions  with the vector heights! 
# Remember you can always review the slides or previous worksheets to see what
# functions we've used before! Or use the ?function_name to try a new one
```{r}
# 10a
class(heights)

# 10b ex. calculating the standard deviation
sd(heights)
```


# Data Frames/tidyverse Review
# Creating your own data frame
# 11a) create 3 different vectors with the length of 6:
# The first vector called boolean_vector containing TRUE or FALSE (boolean data)
# The second vector called favorite_foods containing your favorite foods
# The third vector called my_numbers containing 6, 10, 24, 61, 3, and 200

```{r}
boolean_vector <- c(TRUE, TRUE, FALSE, FALSE, TRUE, TRUE)
favorite_foods <- c("ramen", "gnocchi", "tacos", "dark chocolate", "pho", "horchata")
my_numbers <- c(6, 10, 24, 61, 3, 200)

```

# 11b) put boolean_vector, favorite_foods, and my_numbers into a dataframe
# called my_df. The column names should be boolean, character, and number.
```{r}
my_df = data.frame(boolean = boolean_vector, character = favorite_foods,
                   number = my_numbers)

```

# 11c) look at the first few rows of my_df.
```{r}
head(my_df)
```

# Utilizing the .csv file dataframe
# 12a) read in the pokemon.csv file as the object called pokemon_df
```{r}
pokemon_df <- read.csv("pokemon.csv")
```

# 12b) how many rows and columns are there in pokemon_df?
# 12c) what is the name of the 12th column in the pokemon data frame?
# 12d) what is the name of a column containing character data?
# 12e) what is the name of a column containing numeric or integers?
```{r}
dim(pokemon_df)
# 1196 rows and 13 columns

pokemon_df[12]
# column is called speed

View(pokemon_df)
# columns: Name, Form, Type1, and Type2 contain character data
# columns: Total, HP, Attack, Defense, Sp.. Atk, Sp.. Def, Speed, Generation
# for this key, I will choose the Sp.. Def column

```
# 13a) in your chosen column from 13e (column with numerics or integers)
# are there any NA values?
```{r}
sum(is.na(pokemon_df$Sp..Def))
# yes there are 2 NA values
```

# 13b) compute the following quantities with your chosen column containing numerics or integers: max, min, median, mean, standard deviation
```{r}
pokemon_sp.DEF <- pokemon_df$Sp..Def

# this changes the integer/numeric data to boolean data (true false) and isn't
# what we're looking for
pokemon_sp.DEF_noNA <- !(is.na(pokemon_sp.DEF))

# we can use na.omit() to remove the NAs from the vector
pokemon_sp.DEF_noNA <- na.omit(pokemon_sp.DEF)

# alternatively can use max(pokemon_sp.DEF, na.rm = TRUE)
max_sp.DEF <- max(pokemon_sp.DEF_noNA)
max_sp.DEF

min_sp.DEF <- min(pokemon_sp.DEF_noNA)
min_sp.DEF

mean_sp.DEF <- mean(pokemon_sp.DEF_noNA)
mean_sp.DEF

sd_sp.DEF <- sd(pokemon_sp.DEF_noNA)
sd_sp.DEF
```

# 13c) how many datapoints are less than the median?
# 13d) how many datapoints are more than or equal to the mean?
# 13e) how many datapoints are more than 195 OR less than 80
```{r}
#13c
median_sp.DEF <- median(pokemon_sp.DEF_noNA)
length(pokemon_sp.DEF_noNA[pokemon_sp.DEF_noNA < median_sp.DEF])

# this makes sense because half the data should be less than the median
# 572 datapoints

# 13d
length(pokemon_sp.DEF_noNA[pokemon_sp.DEF_noNA >= mean_sp.DEF])
# 529 datapoints

# 13e 
length(pokemon_sp.DEF_noNA[pokemon_sp.DEF_noNA > 195 | pokemon_sp.DEF_noNA < 80])
# 737 datapoints
```

# 13e) how many datapoints are within 30% of the median?
```{r}
median_30_top <- median_sp.DEF * 1.3
median_30_low <- median_sp.DEF * 0.7

median_30_values <- length(pokemon_sp.DEF_noNA[pokemon_sp.DEF_noNA <= median_30_top & pokemon_sp.DEF_noNA >= median_30_low])

median_30_values
# 707 datapoints
```

# 14a) create a vector called pokemon_Total containing the column called Total in the pokemon_df
# 14b) how many pokemon have a total higher than 450?
# 14c) create a vector called pokemon_Defense containing the column called Defense
# in the pokemon_df.
# challenge question 14d) which pokemon has the most Defense? (hint: use the which() function)
```{r}
pokemon_Total <- pokemon_df$Total

#14b
higher <- length(pokemon_Total[pokemon_Total > 450])
# 638 pokemon

#14c
pokemon_Defense <- pokemon_df$Defense

#14d
pokemon_Defense <- pokemon_df$Defense
max_defense <- max(pokemon_Defense, na.rm = TRUE) # finding max defense stat
index_max_row <- which(pokemon_Defense == 250) # getting the index row for the max defense stat
whos_that_pokemon <- pokemon_df[index_max_row,] # extracting based on the index
whos_that_pokemon

# Eternatus

```

# Debugging: Based on the given code, find the error! This could be using the wrong
# function, spelling, or using the wrong variable etc.! If you're stuck, try running the code 
# what is the error? What do you change to get the expected result?
```{r}
# 15a)  I want to find the max time of my vector called sprint_times. I am expecting 17.10 as the result.
sprint_times <- c(10.55, 17.10, 8.24, 11.45)
max(sprint_time)

# wrong variable name in the function max()
# correct:
max(sprint_times)
```

```{r}
# 15b) I want to determine the median of my vector called years. I am expecting 2002.5 as the result
years <- c(2011, 2024, 2022, 1988, 1994)
mean(years)

# used the function mean() instead of median()
```

```{r}
# 15c) I want to remove the NAs from my vector called route_numbers and calculate the mean.
route_numbers <- s(25, 12, NA, NA, 99, 84)
mean(route_numbers, rm.na = route_numbers))

# A) wrote a non-existent function called s() it should be c() to make a vector
# B) rm.na should be na.rm()
# C) na.rm = TRUE and should not contain a variable
```

Excellent job!! 
Don't forget to save your work! Click on the save icon or (Ctrl/Cmd + S).  



